### 2024-02-01 outputs/paper/paper.qmd
Initiate the project

---
title: "Potential Factors of Decreasing Birth Rates in the US since the Great Recession and Beyond"
subtitle: "Reproduction of 'The Puzzle of Falling US Birth Rates since the Great Recession'(Kearney, Levine & Pardue, 2022)"
author: 
  - Chay Park
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Social-Science-Study."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-02-01 outputs/paper/paper.qmd
Update paper.qmd

@@
 
+#### #### 
+The results section should convey findings.
+The discussion should include three or four sub-sections that each focus on an interesting point, and there should be another sub-section on the weaknesses of your paper, and another on potential next steps for it.
+In the discussion section, and any other relevant section, please be sure to discuss ethics and bias, with reference to relevant literature.
+#### ####
+ data, results, discussion, and references.
 # Introduction

### 2024-02-01 birthrate.qmd
Update and rename outputs/paper/paper.qmd to birthrate.qmd
Update birthrate.qmd
framework of the paper.qmd

--- 2024-02-01 outputs/paper/paper.qmd
@@
 
-You can and should cross-reference sections and sub-sections.
-
-The remainder of this paper is structured as follows. @sec-data....
-
 # Data {#sec-data}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Source
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+## Methodology
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+## Features
 
-Talk more about it.
+### Data Cleaning
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+### Data Modification
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+### Data Visualization
 
 # Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
+The results section should convey findings.
+## Table, graph, table, graph
 
@@
 
-## First discussion point {#sec-first-point}
+In the discussion section, and any other relevant section, please be sure to discuss ethics and bias, with reference to relevant literature.
+
+## Findings
 
@@
 
-## Second discussion point
+## Ethical Implication
 
-## Third discussion point
+## Accounting for bias
 
-## Weaknesses and next steps
+## Limitation
 
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
+## Future Research
 

### 2024-02-03 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Chay Park
+# Date: 3 February 2024 
+# Contact: chay.park@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]

### 2024-02-08 birthrate.qmd
Update birthrate.qmd

@@
 ## Source
+### Trend in US Birth Rates
+
+# Load necessary libraries
+library(ggplot2)
+
+# Read the data from CSV file
+data <- read.csv("fig_1.csv")
+
+# Plotting the data
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_line(color = "blue") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       title = "Birth Rates per 1,000 Women Age 15-44 Over Time",
+       subtitle = "Source: CDC Vital Statistics Births Reports (2015, 2019, 2020)",
+       caption = "Trend in US Birth Rates") +
+  theme_minimal()
 

### 2024-02-08 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -7,7 +7,33 @@
 # Pre-requisites: Need to download the replication package
 
 #### Workspace setup ####
+install.packages(c("knitr", "lubridate", "tidyverse"))
 
+# See a list of available packages
+library(knitr)
+library(lubridate)
+library(tidyverse)
+
+packages <- list_packages(limit = 10)
+packages
 
 #### Clean data ####
 
+# Write CSV file for figure 1
+write_csv(
+  x = trend_us_brate,
+  file = "inputs/data/fig_1.csv"
+)
+head(trend_us_brate)
+
+# Write CSV file for figure 2
+write_csv(
+  x = numbirth,
+  file = "inputs/data/decomp/births_educ_race_age6.csv"
+)
+
+# Write CSV file for figure 3
+write_csv(
+  x = state_birth,
+  file = "inputs/data/map/state_birth_04_19.csv"
+)

### 2024-02-08 scripts/00-simulate_data.R
Update the simulate data

@@ -14,7 +14,6 @@
 library(ggplot2)
 library(dplyr)
 library(tidyr)
-library(usmap)
 
 #### Simulate data ####
 # Simulated birth rate data from 1980 to 2020
@@ -53,10 +52,3 @@ ggplot(subgroup_data, aes(x = year, y = births, color = subgroup)) +
   geom_line() +
   labs(x = "Year", y = "Number of Births", title = "Trends in Birth Rates by Population Subgroup")
 
-# Third table: State-wise Birth Rate Change
-state_data_table <- state_data
-
-# Third figure: Map of US with change in birth rate legend
-usmap::plot_usmap(data = state_data, values = "brate_1544_thsnds_ch_pct", lines = "lightgray") +
-  scale_fill_continuous(name = "Change in Birth Rate", low = "red", high = "green", breaks = c(-15, -10, -5, 0, 5, 10, 15), labels = c("< -10", "-5 to -10", "-5 to 0", "0", "0 to 5", "5 to 10", "> 10")) +
-  labs(title = "Change in Birth Rates by State")

### 2024-02-08 outputs/paper/paper.qmd
Add r scripts for the figure

--- 2024-02-01 outputs/paper/paper.qmd
@@
 
-You can and should cross-reference sections and sub-sections.
-
-The remainder of this paper is structured as follows. @sec-data....
-
 # Data {#sec-data}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+## Source
 
@@
 #| warning: false
-#| message: false
+#| label: fig-usbrate-graph
+#| fig-cap: "Trend in US Birth Rates"
 
-library(rstanarm)
+# Load necessary libraries
+library(ggplot2)
 
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
+
+# Plotting the data
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_line(color = "blue") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
+  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       title = "Trend in US Birth Rates",
+       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+
+```
+```{r}
+# Load necessary libraries
+library(ggplot2)
+library(gridExtra)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))
+
+# Plotting the first line graph (Graph A)
+graph_A <- ggplot(data, aes(x = year)) +
+  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
+  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
+  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
+  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
+  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
+  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
+  scale_color_manual(values = c("Age 15-19" = "blue", "Age 20-24" = "orange", "Age 25-29" = "grey",
+                                 "Age 30-34" = "yellow", "Age 35-39" = "cyan", "Age 40-44" = "green"),
+                     labels = c("Age 15-19", "Age 20-24", "Age 25-29", "Age 30-34", "Age 35-39", "Age 40-44"),
+                     name = "Age Group") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5)) +
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
+  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup",
+       title = "Trend in US Birth Rates by Age Group",
+       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+
+# Plotting the second line graph (Graph B)
+graph_B <- ggplot(data, aes(x = year)) +
+  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
+  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
+  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
+  scale_color_manual(values = c("Hispanic" = "grey", "Black, non-Hispanic" = "orange", "White, non-Hispanic" = "blue"),
+                     labels = c("Hispanic", "Black, non-Hispanic", "White, non-Hispanic"),
+                     name = "Race/Ethnicity") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
+  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
+  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup",
+       title = "Trend in US Birth Rates by Race/Ethnicity",
+       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+
+# Arrange both graphs side by side
+grid.arrange(graph_A, graph_B, nrow = 1)
+
 ```
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+# Load necessary libraries
+library(ggplot2)
+library(gridExtra)
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))
+
+# Plotting the first line graph (Graph A)
+graph_A <- ggplot(data, aes(x = year)) +
+  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
+  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
+  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
+  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
+  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
+  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
+  geom_text(aes(label = "Age 15-19", x = 2020, y = brate_1519), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
+  geom_text(aes(label = "Age 20-24", x = 2020, y = brate_2024), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
+  geom_text(aes(label = "Age 25-29", x = 2020, y = brate_2529), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
+  geom_text(aes(label = "Age 30-34", x = 2020, y = brate_3034), hjust = 0, vjust = -0.5, color = "yellow", family = "Times") +
+  geom_text(aes(label = "Age 35-39", x = 2020, y = brate_3539), hjust = 0, vjust = -0.5, color = "cyan", family = "Times") +
+  geom_text(aes(label = "Age 40-44", x = 2020, y = brate_4044), hjust = 0, vjust = -0.5, color = "green", family = "Times") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) paste(x, "\n")) +
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
+  labs(y = "Births per 1,000 women in relevant population subgroup",
+       title = "A: Five-year age group",
+       caption = "Note: Birth rates by age group, race and ethnicity, and marital status are gathered from CDC Vital Statistics Births Reports.
+Birt rates by Hispanic subpopulation, mother's education are calculated by the original author's NCHS Vital Statistics birth microdata, SEER population data, and the Current Population Survey.") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))
+
+# Plotting the second line graph (Graph B)
+graph_B <- ggplot(data, aes(x = year)) +
+  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
+  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
+  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
+  geom_text(aes(label = "Hispanic", x = 2020, y = brate_hisp), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
+  geom_text(aes(label = "Black, non-Hispanic", x = 2020, y = brate_blacknh), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
+  geom_text(aes(label = "White, non-Hispanic", x = 2020, y = brate_whitenh), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
+  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5), labels = function(x) paste(x, "\n")) +
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
+  labs(title = "B: Race and ethnicity (ages 15-44)") +
+  theme_classic()
+
+# Arrange both graphs side by side
+grid.arrange(graph_A, graph_B, nrow = 1)
+
 ```
+
+## Methodology
+
+## Features
+
+### Data Cleaning
+
+### Data Modification
+
+### Data Visualization
+
+# Results
+The results section should convey findings.
+## Table, graph, table, graph
 

### 2024-02-08 outputs/paper/paper.qmd
Update

@@
 ```
-```{r}
-# Load necessary libraries
-library(ggplot2)
-library(gridExtra)
-
-# Set seed for reproducibility
-set.seed(123)
-
-# Read the data from CSV file
-data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))
-
-# Plotting the first line graph (Graph A)
-graph_A <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
-  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
-  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
-  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
-  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
-  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
-  scale_color_manual(values = c("Age 15-19" = "blue", "Age 20-24" = "orange", "Age 25-29" = "grey",
-                                 "Age 30-34" = "yellow", "Age 35-39" = "cyan", "Age 40-44" = "green"),
-                     labels = c("Age 15-19", "Age 20-24", "Age 25-29", "Age 30-34", "Age 35-39", "Age 40-44"),
-                     name = "Age Group") +
-  geom_vline(xintercept = 2007, linetype = "dotted") +
-  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5)) +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
-  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup",
-       title = "Trend in US Birth Rates by Age Group",
-       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
-  theme_classic()
-
-# Plotting the second line graph (Graph B)
-graph_B <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
-  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
-  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
-  scale_color_manual(values = c("Hispanic" = "grey", "Black, non-Hispanic" = "orange", "White, non-Hispanic" = "blue"),
-                     labels = c("Hispanic", "Black, non-Hispanic", "White, non-Hispanic"),
-                     name = "Race/Ethnicity") +
-  geom_vline(xintercept = 2007, linetype = "dotted") +
-  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
-  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup",
-       title = "Trend in US Birth Rates by Race/Ethnicity",
-       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
-  theme_classic()
-
-# Arrange both graphs side by side
-grid.arrange(graph_A, graph_B, nrow = 1)
-
-```
 
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend-brate-graph
+#| fig-cap: "Trends in Birth Rates by Population Subgroup"
 # Load necessary libraries
@@
 
+```
+
+```{r}
 # Plotting the second line graph (Graph B)
@@
   theme_classic()
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))
 

### 2024-02-08 scripts/02-data_cleaning.R

@@ -1,5 +1,6 @@
 #### Preamble ####
 # Purpose: To sort out relevant necessary data from the data set 
+# ~not sure whether I should haave this script tho
 # Author: Chay Park
 # Date: 4 February 2024
 # Contact: chay.park@mail.utoronto.ca
@@ -13,9 +14,6 @@ install.packages(c("knitr", "lubridate", "tidyverse"))
 library(knitr)
 library(lubridate)
 library(tidyverse)
-
-packages <- list_packages(limit = 10)
-packages
 
 #### Clean data ####
 

### 2024-02-08 outputs/paper/paper.qmd
Update qmd file

@@
   geom_vline(xintercept = 2007, linetype = "dotted") +
-  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  annotate("text", x = 2007, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
   geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
@@
        title = "Trend in US Birth Rates",
-       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.", family = "Times") +
   theme_classic()
@@
 #| message: false
-#| label: fig-trend-brate-graph
+#| label: fig-bratetrend-graph
 #| fig-cap: "Trends in Birth Rates by Population Subgroup"
+
+#### Figure 2 ####
+
+#### ####
+
 # Load necessary libraries
@@
 graph_A <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
-  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
-  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
-  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
-  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
-  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
-  geom_text(aes(label = "Age 15-19", x = 2020, y = brate_1519), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
-  geom_text(aes(label = "Age 20-24", x = 2020, y = brate_2024), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
-  geom_text(aes(label = "Age 25-29", x = 2020, y = brate_2529), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
-  geom_text(aes(label = "Age 30-34", x = 2020, y = brate_3034), hjust = 0, vjust = -0.5, color = "yellow", family = "Times") +
-  geom_text(aes(label = "Age 35-39", x = 2020, y = brate_3539), hjust = 0, vjust = -0.5, color = "cyan", family = "Times") +
-  geom_text(aes(label = "Age 40-44", x = 2020, y = brate_4044), hjust = 0, vjust = -0.5, color = "green", family = "Times") +
+  geom_line(aes(y = brate_1519, color = "blue"), size = 1.0) +
+  geom_line(aes(y = brate_2024, color = "orange"), size = 1.0) +
+  geom_line(aes(y = brate_2529, color = "grey"), size = 1.0) +
+  geom_line(aes(y = brate_3034, color = "yellow"), size = 1.0) +
+  geom_line(aes(y = brate_3539, color = "cyan"), size = 1.0) +
+  geom_line(aes(y = brate_4044, color = "green"), size = 1.0) +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_1519, label = "Age 15-19"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2024, label = "Age 20-24"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2529, label = "Age 25-29"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3034, label = "Age 30-34"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3539, label = "Age 35-39"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_4044, label = "Age 40-44"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
   geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
   geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) paste(x, "\n")) +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
-  labs(y = "Births per 1,000 women in relevant population subgroup",
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = NULL, 
+       y = 'Births per 1,000 women in relevant population subgroup',
        title = "A: Five-year age group",
-       caption = "Note: Birth rates by age group, race and ethnicity, and marital status are gathered from CDC Vital Statistics Births Reports.
-Birt rates by Hispanic subpopulation, mother's education are calculated by the original author's NCHS Vital Statistics birth microdata, SEER population data, and the Current Population Survey.") +
+       caption = NULL,
+       family = "Times") +
   theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))
+  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times"),
+        plot.title = element_text(size = 10, family = "Times"),
+        legend.position = "none")
 
-```
-
-```{r}
 # Plotting the second line graph (Graph B)
 graph_B <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
-  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
-  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
-  geom_text(aes(label = "Hispanic", x = 2020, y = brate_hisp), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
-  geom_text(aes(label = "Black, non-Hispanic", x = 2020, y = brate_blacknh), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
-  geom_text(aes(label = "White, non-Hispanic", x = 2020, y = brate_whitenh), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
-  geom_vline(xintercept = 2007, linetype = "dotted") +
-  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5), labels = function(x) paste(x, "\n")) +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
-  labs(title = "B: Race and ethnicity (ages 15-44)") +
-  theme_classic()
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))
+  geom_line(aes(y = brate_hisp, color = "grey"), size = 1.2) +
+  geom_line(aes(y = brate_blacknh, color = "orange"), size = 1.2) +
+  geom_line(aes(y = brate_whitenh, color = "blue"), size = 1.2) +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") + scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) + scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) + labs(x = NULL, y = NULL,
+       title = "B: Race and ethnicity (ages 15-44)",
+       caption = "Birth rates by age group, race, and ethnicity, and marital status gathered from CDC Vital Statistics Births Reports.
+Birth rates by Hispanic subpopulation, mother's education are calculated by the original author's using NCHS Vital Statistics birth micro data, SEER population data, and the Current Population Survey.") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times"),
+        plot.title = element_text(size = 10, family = "Times"),
+        legend.position = "none")
 

### 2024-02-09 outputs/paper/paper.qmd
update

--- 2024-02-08 outputs/paper/paper.qmd
@@
 #| warning: false
-#| label: fig-usbrate-graph
-#| fig-cap: "Trend in US Birth Rates"
-
-# Load necessary libraries
-library(ggplot2)
-
-# Set seed for reproducibility
-set.seed(123)
-
-# Read the data from CSV file
-data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
-
-# Plotting the data
-ggplot(data, aes(x = year, y = brate_all)) +
-  geom_line(color = "blue") +
-  geom_vline(xintercept = 2007, linetype = "dotted") +
-  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
-  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
-  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
-  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
-  labs(x = "Year", y = "Births per 1,000 women age 15-44",
-       title = "Trend in US Birth Rates",
-       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
-  theme_classic()
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
 #| message: false
-#| label: fig-trend-brate-graph
+#| label: fig-bratetrend-graph
 #| fig-cap: "Trends in Birth Rates by Population Subgroup"
@@
 graph_A <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
-  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
-  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
-  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
-  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
-  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
-  geom_text(aes(label = "Age 15-19", x = 2020, y = brate_1519), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
-  geom_text(aes(label = "Age 20-24", x = 2020, y = brate_2024), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
-  geom_text(aes(label = "Age 25-29", x = 2020, y = brate_2529), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
-  geom_text(aes(label = "Age 30-34", x = 2020, y = brate_3034), hjust = 0, vjust = -0.5, color = "yellow", family = "Times") +
-  geom_text(aes(label = "Age 35-39", x = 2020, y = brate_3539), hjust = 0, vjust = -0.5, color = "cyan", family = "Times") +
-  geom_text(aes(label = "Age 40-44", x = 2020, y = brate_4044), hjust = 0, vjust = -0.5, color = "green", family = "Times") +
+  geom_line(aes(y = brate_1519, color = "blue"), size = 0.4) +
+  geom_line(aes(y = brate_2024, color = "orange"), size = 0.4) +
+  geom_line(aes(y = brate_2529, color = "grey"), size = 0.4) +
+  geom_line(aes(y = brate_3034, color = "yellow"), size = 0.4) +
+  geom_line(aes(y = brate_3539, color = "cyan"), size = 0.4) +
+  geom_line(aes(y = brate_4044, color = "green"), size = 0.4) +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_1519, label = "Age 15-19"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2024, label = "Age 20-24"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2529, label = "Age 25-29"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3034, label = "Age 30-34"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3539, label = "Age 35-39"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_4044, label = "Age 40-44"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
   geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
   geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) paste(x, "\n")) +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
-  labs(y = "Births per 1,000 women in relevant population subgroup",
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = NULL, 
+       y = 'Births per 1,000 women in relevant population subgroup',
        title = "A: Five-year age group",
-       caption = "Note: Birth rates by age group, race and ethnicity, and marital status are gathered from CDC Vital Statistics Births Reports.
-Birt rates by Hispanic subpopulation, mother's education are calculated by the original author's NCHS Vital Statistics birth microdata, SEER population data, and the Current Population Survey.") +
+       caption = NULL,
+       family = "Times") +
   theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))
+  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 5),
+        plot.title = element_text(size = 7, family = "Times"),
+        legend.position = "none")
+
+# Plotting the second line graph (Graph B)
+graph_B <- ggplot(data, aes(x = year)) +
+  geom_line(aes(y = brate_hisp, color = "grey"), size = 0.6) +
+  geom_line(aes(y = brate_blacknh, color = "orange"), size = 0.6) +
+  geom_line(aes(y = brate_whitenh, color = "blue"), size = 0.6) +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") + scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) + scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) + labs(x = NULL, y = NULL,
+       title = "B: Race and ethnicity (ages 15-44)",
+       caption = "Birth rates by age group, race, and ethnicity, and marital status gathered from CDC Vital Statistics Births Reports.
+Birth rates by Hispanic subpopulation, mother's education are calculated by the original author's using NCHS Vital Statistics birth micro data, SEER population data, and the Current Population Survey.") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 5),
+        plot.title = element_text(size = 7, family = "Times"),
+        legend.position = "none")
+
+# Arrange both graphs side by side
+grid.arrange(graph_A, graph_B, nrow = 1)
 
@@
 ```{r}
-# Plotting the second line graph (Graph B)
-graph_B <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
-  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
-  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
-  geom_text(aes(label = "Hispanic", x = 2020, y = brate_hisp), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
-  geom_text(aes(label = "Black, non-Hispanic", x = 2020, y = brate_blacknh), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
-  geom_text(aes(label = "White, non-Hispanic", x = 2020, y = brate_whitenh), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
-  geom_vline(xintercept = 2007, linetype = "dotted") +
-  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5), labels = function(x) paste(x, "\n")) +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
-  labs(title = "B: Race and ethnicity (ages 15-44)") +
-  theme_classic()
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))
-
-# Arrange both graphs side by side
-grid.arrange(graph_A, graph_B, nrow = 1)
 

### 2024-02-09 outputs/paper/paper.qmd
Update

@@
   geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) +
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = seq(0, 140, by = 20)) +
   scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
@@
 graph_B <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = brate_hisp, color = "grey"), size = 0.6) +
-  geom_line(aes(y = brate_blacknh, color = "orange"), size = 0.6) +
-  geom_line(aes(y = brate_whitenh, color = "blue"), size = 0.6) +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
-  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") + scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) + scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) + labs(x = NULL, y = NULL,
+  geom_line(aes(y = brate_hisp, color = "grey"), size = 0.4) +
+  geom_line(aes(y = brate_blacknh, color = "orange"), size = 0.4) +
+  geom_line(aes(y = brate_whitenh, color = "blue"), size = 0.4) +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), size = 2, family = "Times") +
+  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2007, y = 120, label = "2007", color = "black", family = "Times") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") + scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = seq(0, 140, by = 20)) + scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) + labs(x = NULL, y = NULL,
        title = "B: Race and ethnicity (ages 15-44)",
@@
 
-# Arrange both graphs side by side
-grid.arrange(graph_A, graph_B, nrow = 1)
+# Read the data from the new CSV file fig_2e
+data_2e <- read.csv(file = here::here("outputs/data/fig_2e.csv"))
 
-```
+# Plotting the third line graph (Graph C)
+graph_C <- ggplot(data_2e, aes(x = year)) +
+  geom_line(aes(y = brate_unmarried, color = "blue"), size = 0.4) +
+  geom_line(aes(y = brate_married, color = "orange"), size = 0.4) +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
+  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
+  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = seq(0, 140, by = 20)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = NULL, 
+       y = 'Births per 1,000 women in relevant population subgroup',
+       title = "C: Marital Status",
+       caption = NULL,
+       family = "Times") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 5),
+        plot.title = element_text(size = 7, family = "Times"),
+        legend.position = "none")
 
-```{r}
+# divide frame in 3x1 grid
+par(mfrow = c(3,1))
+
+# draw 3 graphs
+plot(graph_A)
+plot(graph_B)
+plot(graph_C)
 

### 2024-02-09 outputs/paper/paper.qmd
Update

@@
 ## Source
-
 ```{r}
@@
 #| label: fig-bratetrend-graph
+#| fig-cap: "Trend in US Birth Rates"
+
+# Load necessary libraries
+library(ggplot2)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
+
+# Plotting the data
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_line(color = "blue") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
+  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       title = "Trend in US Birth Rates",
+       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+```
+
+# 낙태법 시행일, 법적으로 진행된 일, economic crisis other than great recession 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-bratepop-graph
 #| fig-cap: "Trends in Birth Rates by Population Subgroup"
+
 # Load necessary libraries
@@
 ```
+# 낙태법 적용 나이, 히스패닉과 아프리칸, 서양인들의 종교적 문화 차이-whether they accept the abortion or not, cultural perspectives, whether marriage affects the decision of keeping a child or not.<- financial reason
+
+# map
 

### 2024-02-09 outputs/paper/paper.qmd
Update paper.qmd

--- 2024-02-09 outputs/paper/paper.qmd
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Decreasing birth rates became the conversation topic globally in recent years.
+Kearney, Levine, and Pardue (2022) analyzed possible factors in demographical, economical and social aspects from the period 1980 to 2020.
+By reproducing some applications of the 'The Puzzle of Falling US Birth Rates Since the Great Recession,' this paper demonstrate potential factors correlating aspects on economics, religions, laws and enforcement applied in each states."
 format: pdf
@@
 # Introduction
+1) broader context to motivate
+2) some detail about what the paper is about
+3) a clear gap that needs to be filled
+4) what was done
+5) what was found
+6) why it is important
+7) the structure of the paper
 
+In an economical aspect, birth rates are important factor that decides the growth of a country. 
+The paper is about the factors that was not discussed in the Kearney's paper.
+They pointed out that as of 2007, the Great Recession, was the turning point of declining tendency in birth rates, however, it is beyond just demographical, economical, social and political factors.
+In this paper we will
+
+Over the 40 years,
+
+Age over 30
+
+* what was done, why and what was found, 3-4 paragraphs
+
+From now on, every white refers to non-Hispanic, caucassian. Every black refers to non-Hispanic, African American. Hispanic refers to people with Hispanic ethinicity majority from south America. 
 # Data {#sec-data}
@@
 ## Source
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-bratetrend-graph
+#| fig-cap: "Trend in US Birth Rates"
+
+# Load necessary libraries
+library(ggplot2)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
+
+# Plotting the data
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_line(color = "blue") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
+  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       title = "Trend in US Birth Rates",
+       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+```
+Figure 1 Trend in US Birth Rates, not just the Great Recession was there another events? 
+낙태법 시행일, 법적으로 진행된 일, economic crisis other than great recession
 
@@
 #| message: false
-#| label: fig-bratetrend-graph
+#| label: fig-bratepop-graph
 #| fig-cap: "Trends in Birth Rates by Population Subgroup"
+
 # Load necessary libraries
@@
 
-# Read the data from the new CSV file fig_2e
-data_2e <- read.csv(file = here::here("outputs/data/fig_2e.csv"))
-
-# Plotting the third line graph (Graph C)
-graph_C <- ggplot(data_2e, aes(x = year)) +
-  geom_line(aes(y = brate_unmarried, color = "blue"), size = 0.4) +
-  geom_line(aes(y = brate_married, color = "orange"), size = 0.4) +
-  geom_vline(xintercept = 2007, linetype = "dotted") +
-  annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
-  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
-  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = seq(0, 140, by = 20)) +
-  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
-  labs(x = NULL, 
-       y = 'Births per 1,000 women in relevant population subgroup',
-       title = "C: Marital Status",
-       caption = NULL,
-       family = "Times") +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 5),
-        plot.title = element_text(size = 7, family = "Times"),
-        legend.position = "none")
-
-# divide frame in 3x1 grid
-par(mfrow = c(3,1))
-
-# draw 3 graphs
-plot(graph_A)
-plot(graph_B)
-plot(graph_C)
+# Arrange both graphs side by side
+grid.arrange(graph_A, graph_B, nrow = 1)
 
 ```
+Figure 2 A, 낙태법 적용 나이, 기술의 발전으로 인한 낙태 도입률, -> 경제적으로 안정되거나 아이의 부담을 지기 싫은 젊은 여성들은 낙태 ㅗ 기술의 발전으로 인한 건강 유지 및 출산 정책으로 인한 30세 이상의 경제적으로 안정된 여성들은 high willingness to give births, showing ascending pattern in birth rates over the 40 years.
+
+Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity
+
+# 낙태법 적용 나이, 히스패닉과 아프리칸, 서양인들의 종교적 문화 차이-whether they accept the abortion or not, cultural perspectives, whether marriage affects the decision of keeping a child or not.<- financial reason
+
+# map
 

### 2024-02-09 outputs/paper/paper.qmd
Update paper.qmd

@@
 5) what was found
-6) why it is important
-7) the structure of the paper
+6) why it is important - replacement level fertility
+7) the structure of the paper - 1. trend in US Birth Rates 2. Trends in Birth Rates by Population Subgroup 2.1 Five-year age group 2.2 Race and ethinicity
 
@@
 From now on, every white refers to non-Hispanic, caucassian. Every black refers to non-Hispanic, African American. Hispanic refers to people with Hispanic ethinicity majority from south America.
+
 # Data {#sec-data}
+
+## Methodology
+The data for the trend in US Birth Rates, births per 1,000 women age 15-44 from the period 1980 to 2020 was acquired from CDC Vital Statistics Births Reports for 2015, 2019 and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
+Another trends in Birth Rates by Population Subgroup, specifically birth rates by age group, race and ethnicity are again gathered from CDC Vital Statistics Births Reports.
+Original paper uses single-age population counts, among all races from 1969-2019 and by race and Hispanic origin from 1990-2019, from the CDC SEER database (CDC NCI 2021).
+
+## Features
+
+# Results
+The results section should convey findings.
+## Table, graph, table, graph
 
@@
 
-# map
-
-## Methodology
-
-## Features
-
-### Data Cleaning
-
-### Data Modification
-
-### Data Visualization
-
-# Results
-The results section should convey findings.
-## Table, graph, table, graph
+Map
 

### 2024-02-09 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-Age over 30
-
 * what was done, why and what was found, 3-4 paragraphs
 
-From now on, every white refers to non-Hispanic, caucassian. Every black refers to non-Hispanic, African American. Hispanic refers to people with Hispanic ethinicity majority from south America.
+We replicate the paper by Kearney, Levine, and Pardue (2022) with a focus on the following research questions: 
+- What is the trend in the US birth rates over the period 1980 to 2020?
+- How does the birth rates vary by the age group?
+- How does the birth rates vary by the race and ethnicity of mothers'?
+
+The original code in the replication package is written in Stata.
+However, this paper utilizes R (R Core Team 2020) to analyze a dataset from the replication package.
+Packages such as ggplot2 (Wickham et al.
+2016) for creating graphical representations, kableExtra (Zhu et al.
+2021) for enhancing table aesthetics, and lubridate (Grolemund et al.
+2021) for handling date-related operations were employed.
+Additionally, the Tidyverse framework (Wickham et al.
+2019) ensures cohesiveness, while Dplyr (Wickham et al.
+2021) facilitates data manipulation.
 
@@
 ## Features
+From now on, every white refers to non-Hispanic, caucassian. Every black refers to non-Hispanic, African American. Hispanic refers to people with Hispanic ethinicity majority from south America.
 

### 2024-02-09 outputs/paper/paper.qmd
Update paper.qmd

@@
 Original paper uses single-age population counts, among all races from 1969-2019 and by race and Hispanic origin from 1990-2019, from the CDC SEER database (CDC NCI 2021).
+Data is available freely at https://seer.cancer.gov/popdata/download.html and raw files can be found in the data/pop/folder of the original replication package.
 
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-race-table
+#| tbl-cap: "Race and ethinicity (ages 15-44)"
+
+#### Table 1a ####
+# Count the number of births per 1,000 women in relevant population subgroup by their race and ethinicity in the ages cohort from 15 to 44. 
+#### ####
+
+# Install the necessary packages if not already installed
+install.packages(c("knitr","lubridate","tidyverse","kableExtra"))
+
+# Load required libraries
+library(dplyr)
+library(knitr)
+library(lubridate)
+library(tidyverse)
+library(kableExtra)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Load the CSV file
+
+```
 Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity

### 2024-02-09 outputs/paper/paper.qmd
update

--- 2024-02-09 outputs/paper/paper.qmd
@@
 
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))
+
+# Filter the data to include years starting from 1990
+filtered_data <- data %>%
+  filter(year >= 1990)
+
+# Select the required columns and rename them
+selected_data <- filtered_data %>%
+  select(Year = year, White = brate_whitenh, Black = brate_blacknh, Hispanic = brate_hisp)
+
+# Print the table using additional styling
+kable(selected_data, caption = "Trends in Birth Rates by Population Race and Ethnicity (ages 15-44)", format = "html") %>%
+  kable_styling(full_width = FALSE) %>%
+  add_header_above(c("","Births per 1,000 women" = 3))
+
+```
+
 Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity

### 2024-02-10 outputs/paper/paper.qmd
Update

@@
 bibliography: references.bib
+prefer-html: true
 ---
@@
 We replicate the paper by Kearney, Levine, and Pardue (2022) with a focus on the following research questions:
-- What is the trend in the US birth rates over the period 1980 to 2020?
-- How does the birth rates vary by the age group?
-- How does the birth rates vary by the race and ethnicity of mothers'?
+
+* What is the trend in the US birth rates over the period 1980 to 2020?
+
+* How does the birth rates vary by the age group?
+
+* How does the birth rates vary by the race and ethnicity of mothers'?
 
@@
   labs(x = "Year", y = "Births per 1,000 women age 15-44",
-       title = "Trend in US Birth Rates",
        caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
@@
 ```
+
 Figure 1 Trend in US Birth Rates, not just the Great Recession was there another events? 
@@
   theme_classic() +
-  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 5),
-        plot.title = element_text(size = 7, family = "Times"),
+  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 8),
+        plot.title = element_text(size = 10, family = "Times"),
         legend.position = "none")
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-trendssub-table
+#| tbl-cap: "Trends in Birth Rates by Population Race and Ethnicity"
+
 # Load necessary libraries
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trendssub-graph
+#| fig-cap: "Trends in Birth Rates by Population Race and Ethnicity"
+
+# Load required libraries
+library(dplyr)
+library(ggplot2)
+
+# Refer to the table a
+# Assume selected_data contains the relevant data
+
+# Create bar graphs for each racial/ethnic group
+bar_plot <- ggplot(selected_data, aes(x = Year)) +
+  geom_bar(aes(y = Hispanic, fill = "Hispanic"), stat = "identity", position = "stack") +
+  geom_bar(aes(y = Black, fill = "Black"), stat = "identity", position = "stack") +
+  geom_bar(aes(y = White, fill = "White"), stat = "identity", position = "stack") +
+  labs(x = "Year",
+       y = "Births per 1,000 women relevant population subgroup") +
+  scale_fill_manual(values = c(Hispanic = "grey", Black = "orange", White = "blue"),
+                    name = "Race and ethnicity") +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(1990, 2020, by = 5), minor_breaks = seq(1990, 2020, by = 1)) +
+  ylim(0, 120)
+
+# Print the bar plot
+print(bar_plot)
+```
 Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity

### 2024-02-10 outputs/paper/paper.qmd
add some writings structure

@@
 1) broader context to motivate
+According to dictionary, birth rate is defined as " ."
+
 2) some detail about what the paper is about
@@
   geom_vline(xintercept = 2007, linetype = "dotted") +
-  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  annotate("text", x = 2009, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
   geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
@@
   labs(x = "Year", y = "Births per 1,000 women age 15-44",
-       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
   theme_classic()
@@
 Figure 1 Trend in US Birth Rates, not just the Great Recession was there another events? 
-낙태법 시행일, 법적으로 진행된 일, economic crisis other than great recession
+abortion law, law enforcement dates, economic crisis other than great recession
 
@@
   geom_line(aes(y = brate_4044, color = "green"), size = 0.4) +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_1519, label = "Age 15-19"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2024, label = "Age 20-24"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2529, label = "Age 25-29"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3034, label = "Age 30-34"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3539, label = "Age 35-39"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_4044, label = "Age 40-44"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_1519, label = "Age 15-19"), x = 1980, y = 60, hjust = -0.1, vjust = -0.5, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2024, label = "Age 20-24"), x = 1980, y = 100, hjust = -0.1, vjust = -0.5, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2529, label = "Age 25-29"), x = 1980, y = 125 , hjust = -0.1, vjust = -0.5, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3034, label = "Age 30-34"), x = 1980, y = 80, hjust = -0.1, vjust = -0.5, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3539, label = "Age 35-39"), x = 1980, y = 25, hjust = -0.1, vjust = -0.5, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_4044, label = "Age 40-44"), x = 1980, y = 5, hjust = -0.1, vjust = -0.5, size = 3, family = "Times") +
   geom_vline(xintercept = 2007, linetype = "dotted") +
-  annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
+  annotate("text", x = 2011, y = 122, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
   geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
@@
   theme_classic() +
-  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 8),
+  theme(axis.text.x = element_text(angle = -45, family = "Times", size = 8),
         plot.title = element_text(size = 10, family = "Times"),
@@
   geom_line(aes(y = brate_whitenh, color = "blue"), size = 0.4) +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), size = 2, family = "Times") +
-  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), size = 2, family = "Times") +
-  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2007, y = 120, label = "2007", color = "black", family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), x = 1996, y = 110, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), x = 1998, y = 90, size = 3, family = "Times") +
+  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), x = 1998, y = 50, size = 3, family = "Times") +
+  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2010, y = 125, label = "2007", color = "black", family = "Times") +
   geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") + scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = seq(0, 140, by = 20)) + scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) + labs(x = NULL, y = NULL,
-       title = "B: Race and ethnicity (ages 15-44)",
-       caption = "Birth rates by age group, race, and ethnicity, and marital status gathered from CDC Vital Statistics Births Reports.
-Birth rates by Hispanic subpopulation, mother's education are calculated by the original author's using NCHS Vital Statistics birth micro data, SEER population data, and the Current Population Survey.") +
+       title = "B: Race and ethnicity (ages 15-44)") +
   theme_classic() +
-  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times", size = 8),
+  theme(axis.text.x = element_text(angle = -45, family = "Times", size = 8),
         plot.title = element_text(size = 10, family = "Times"),
@@
 ```
-Figure 2 A, 낙태법 적용 나이, 기술의 발전으로 인한 낙태 도입률, -> 경제적으로 안정되거나 아이의 부담을 지기 싫은 젊은 여성들은 낙태 ㅗ 기술의 발전으로 인한 건강 유지 및 출산 정책으로 인한 30세 이상의 경제적으로 안정된 여성들은 high willingness to give births, showing ascending pattern in birth rates over the 40 years.
+Figure 2 A, abortion law applicable age ~,
+
+Technology enhancement in medical field allows the abortion to be held more easily compared to 1980, 1990s.
+This change allow the young women who can afford the abortion or vice versa who can't afford giving birth to get the abortion.
+In the meantime, due to the advancement of technology, women at relatively old age can give the birth in their late 30s or even 40s.
+As well as change in child birth policy and health maintenance was eased compared to the old days in 1980s, 1990s, that promotes women at age older than 30 who are relatively financially stable have high willingness to give births, showing ascending pattern in birth rates over the 40 years.
 
@@
 ```
-Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity
+Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity.
 
-# 낙태법 적용 나이, 히스패닉과 아프리칸, 서양인들의 종교적 문화 차이-whether they accept the abortion or not, cultural perspectives, whether marriage affects the decision of keeping a child or not.<- financial reason
+Here we are dissecting the cultural and religious difference between the races among the population subgroup. 
+Hispanic, majority catholic,
+White, christian, catholic, etc
+Black, christian, catholic, etc
 
-Map
+It is prevailed that certain religion does not allow the abortion, saying that it disobey the words of God or whatever.
+
+-whether they accept the abortion or not, cultural perspectives, whether marriage affects the decision of keeping a child or not.<- financial reason
+
+Might include Map or not depending on the sources and references I find, see how long it'll gonna be.
 
@@
 ## Limitation
+As mentioned in the original paper, the population subgroup can be diverged as they mix over the time.
+It is hard to sort out the race completely once they start to mixing it up.
+Therefore, over the time, we don't know the population subgroup Hispanic, Black, White are purely Hispanic, Black, White throughout the period.
+
+Also, we should not disregard the fact that for all statistical data, they always going to be a missing value that is not properly collected.
+A lot of people might not just report or illegally keep the population due to their illegal status in the US.
+
+Therefore, the data might be skewed due to aforementioned reasons.
 
 ## Future Research
+It might be interesting to expand the research worldwide, not just limiting ourselves into the US data.
+It is broadly known that Chinese and Indian population are growing exponentially while the birth rate in other developed countries continue to grow small.
+It might be interesting to research more what's behind.
 

### 2024-02-10 scripts/02-data_cleaning.R
updat3e

@@ -17,21 +17,3 @@ library(tidyverse)
 
 #### Clean data ####
 
-# Write CSV file for figure 1
-write_csv(
-  x = trend_us_brate,
-  file = "inputs/data/fig_1.csv"
-)
-head(trend_us_brate)
-
-# Write CSV file for figure 2
-write_csv(
-  x = numbirth,
-  file = "inputs/data/decomp/births_educ_race_age6.csv"
-)
-
-# Write CSV file for figure 3
-write_csv(
-  x = state_birth,
-  file = "inputs/data/map/state_birth_04_19.csv"
-)

### 2024-02-11 outputs/paper/paper.qmd
Update the reference
Update paper.qmd

@@
 1) broader context to motivate
-According to dictionary, birth rate is defined as " ."
-
 2) some detail about what the paper is about
@@
 Over the 40 years,
+
+According to the Centers for Disease Control and Prevention, birth rate is defined as "the number of live births per 1,000 population." It is calculated by dividing the number of live births in a population in a year by the midyear resident population.
+More specifically, for census years, rates are based on unrounded census counts of the resident population as of April 1.
+For the noncensus years 1981 - 1989, rates are based on the midyear resident population, rounded to thousands (National Center for Health Statistics 2023).
+In an economical aspect, birth rates are an important factor that decides the growth of a country.
+In recent years, low birth rate became a conversation topic both nation-wide and worldwide.
+Many countries are rising this issue of declining birth rates and posing alternative measures to compensate the population drop.
+For instance,
 
@@
 ## Limitation
+Beginning in 1997, the birth rate for the maternal age group 45-49 includes data for mothers aged 45 and over in the numerator and is based on the population of women aged 45-49 in the denominator (National Center for Health Statistics 2023).
+However as the dataset is missing for the population of women aged 45-49 before 1997 and the paper counts the data from the period 1980 to 2020, the women population in the age group 45-49 is eliminated.
+
 As mentioned in the original paper, the population subgroup can be diverged as they mix over the time.
@@
 # References
+
+---
+nocite: '@*'
+---

### 2024-02-11 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-In an economical aspect, birth rates are important factor that decides the growth of a country. 
-The paper is about the factors that was not discussed in the Kearney's paper.
-They pointed out that as of 2007, the Great Recession, was the turning point of declining tendency in birth rates, however, it is beyond just demographical, economical, social and political factors.
-In this paper we will
-
-Over the 40 years,
-
 According to the Centers for Disease Control and Prevention, birth rate is defined as "the number of live births per 1,000 population." It is calculated by dividing the number of live births in a population in a year by the midyear resident population.
@@
 Many countries are rising this issue of declining birth rates and posing alternative measures to compensate the population drop.
-For instance,
+For instance, Germany birth rate (births per 1000 people) in 2024 hits 9.3 which is lower than Canada 10.0 and the US 12.0 ().
+Germany takes an action to actively recruiting skilled workers from abroad and reforming its citizenship law by allowing dual citizenship
 

### 2024-02-11 outputs/paper/paper.qmd
Update paper.qmd

@@
 # Introduction
-1) broader context to motivate
-2) some detail about what the paper is about
-3) a clear gap that needs to be filled
-4) what was done
-5) what was found
-6) why it is important - replacement level fertility
-7) the structure of the paper - 1. trend in US Birth Rates 2. Trends in Birth Rates by Population Subgroup 2.1 Five-year age group 2.2 Race and ethinicity
 
@@
 Many countries are rising this issue of declining birth rates and posing alternative measures to compensate the population drop.
-For instance, Germany birth rate (births per 1000 people) in 2024 hits 9.3 which is lower than Canada 10.0 and the US 12.0 ().
-Germany takes an action to actively recruiting skilled workers from abroad and reforming its citizenship law by allowing dual citizenship
+For instance, Germany birth rate (births per 1000 people) in 2024 hits 9.3 which is lower than Canada 10.0 and the US 12.0 (Macrotrends 2024).
+Germany takes an action to actively recruiting skilled workers from abroad (Grunau 2023) and reforming its citizenship law by allowing dual citizenship (Knight 2024) as some measures to the declining birth rates and overall population.
+Last but not least, birth rate also indicates the replacement level fertility meaning that the level of fertility at which a population exactly replaces itself from one generation to the next.
+In developed countries, replacement level fertility can be taken as requiring an average of 2.1 children per woman (Craig 1994).
+Thus, it is a matter of time that birth rates becomes a key topic globally in mid to long term.
 
-* what was done, why and what was found, 3-4 paragraphs
+Two economists, Melissa S.Kearney, Phillip B.Levine and a PhD Candidate in Economics, Luke Pardue wrote a paper about the dramatic decline in birth rates in the United States between 2007 and 2020 and suggest possible reasonings.
+They presented the overall reduction in the birth rate among the population groups of women, including teens, Hispanic women, and college-educated white women.
+They suggested the Great Recession was the main reasoning of the decline in the early part of the period between 2007 and 2020, however they were unable to identify any other economic, policy, or social factor that has changed since 2007 which support the decline beyond the Great Recession (Kearney, Levine, and Pardue 2022).
+The purpose of this paper is to replicate the figures and the tables from the original paper and be able to find out further reasonings of declining birth rates in the US.
+Beyond the Great Recession, this paper diverge the perspectives in population subgroups and dissect more in five-year age group and race and ethnicity group in depth.
+
+This paper used a replication package of the original and utilized cleaned data CSV file to reproduce the related figures and tables.
+In which allow to draw the correlation of the data set with political, economical, social policy measures that was not covered in the original paper.
+Furthermore, this paper investigates some socio-cultural aspects including sexual education, marriage population, financial stability of the elder women at the age of over 30, and the technology in healthcare.
+-what was found-
+
+First, the paper showcases the trend in the US Birth Rates and discuss the relevant factors along with the figure.
+Then it demonstrate trends in birth rates by population subgroup.
+The focused population subgroups are 1.Five-year age group 2 Race and ethinicity.
+Along with the graphs and tables, the paper correlates various aspects including education, socio-economic, finance and take them into account to explain the declining birth rates.
 

### 2024-02-11 outputs/paper/paper.qmd
Update paper.qmd
check grammar and tone

--- 2024-02-11 outputs/paper/paper.qmd
@@
 # Introduction
-1) broader context to motivate
-2) some detail about what the paper is about
-3) a clear gap that needs to be filled
-4) what was done
-5) what was found
-6) why it is important - replacement level fertility
-7) the structure of the paper - 1. trend in US Birth Rates 2. Trends in Birth Rates by Population Subgroup 2.1 Five-year age group 2.2 Race and ethinicity
 
-According to the Centers for Disease Control and Prevention, birth rate is defined as "the number of live births per 1,000 population." It is calculated by dividing the number of live births in a population in a year by the midyear resident population.
-More specifically, for census years, rates are based on unrounded census counts of the resident population as of April 1.
-For the noncensus years 1981 - 1989, rates are based on the midyear resident population, rounded to thousands (National Center for Health Statistics 2023).
-In an economical aspect, birth rates are an important factor that decides the growth of a country.
-In recent years, low birth rate became a conversation topic both nation-wide and worldwide.
-Many countries are rising this issue of declining birth rates and posing alternative measures to compensate the population drop.
-For instance, Germany birth rate (births per 1000 people) in 2024 hits 9.3 which is lower than Canada 10.0 and the US 12.0 ().
-Germany takes an action to actively recruiting skilled workers from abroad and reforming its citizenship law by allowing dual citizenship
+According to the Centers for Disease Control and Prevention, the birth rate is defined as "the number of live births per 1,000 population." This rate is calculated by dividing the number of live births in a population in a year by the midyear resident population.
+Specifically, for census years, rates are calculated based on the unrounded census counts of the resident population as of April 1st.
+From an economic perspective, birth rates are an important factor that determines the growth of a country.
+In recent years, the low birth rate has become a conversation topic both nationwide and worldwide.
+Many countries are addressing this issue of declining birth rates and proposing alternative measures to compensate for the population drop.
+For instance, Germany's birth rate (births per 1000 people) in 2024 is 9.3, which is lower than Canada's 10.0 and the US's 12.0 (Macrotrends 2024).
+Germany is taking action by actively recruiting skilled workers from abroad (Grunau 2023) and reforming its citizenship law to allow dual citizenship (Knight 2024) as some measures to combat declining birth rates and overall population.
+Finally, the birth rate also serves as an indicator of replacement level fertility, which in developed countries can be taken as requiring an average of 2.1 children per woman (Craig 1994).
+Therefore, birth rates are poised to become a key global topic in the mid to long term.
 
-* what was done, why and what was found, 3-4 paragraphs
+Two economists, Melissa S.
+Kearney and Phillip B.
+Levine, along with Luke Pardue, a PhD Candidate in Economics, authored a paper addressing the dramatic decline in birth rates in the United States between 2007 and 2020 and proposed possible explanations.
+They examined the overall reduction in the birth rate among various population groups of women, including teens, Hispanic women, and college-educated white women.
+While they attributed the main cause of the decline in the early part of the period to the Great Recession, they were unable to identify any other economic, policy, or social factors that have changed since 2007 to support the decline beyond the Great Recession (Kearney, Levine, and Pardue 2022).
+The objective of this paper is to replicate the figures and tables from the original study and to explore additional factors contributing to the declining birth rates in the US.
+In addition to examining the impact of the Great Recession, this paper delves into the perspectives of population subgroups, analyzing data in five-year age groups and by race and ethnicity to gain deeper insights.
 
-We replicate the paper by Kearney, Levine, and Pardue (2022) with a focus on the following research questions:
+This paper utilized a replication package of the original study and employed a cleaned data CSV file to reproduce the related figures and tables.
+This approach enabled us to draw correlations between the dataset and political, economic, and social policy measures that were not covered in the original paper.
+Furthermore, this paper investigates various socio-cultural aspects, including sexual education, marriage rates, financial stability among women over the age of 30, and the role of technology in healthcare.
 
-* What is the trend in the US birth rates over the period 1980 to 2020?
+-what was found-
 
-* How does the birth rates vary by the age group?
+First, the paper showcases the trend in US birth rates and discusses the relevant factors along with the figure.
+It then demonstrates trends in birth rates by population subgroup, focusing on two key areas: five-year age groups and race and ethnicity.
+Alongside the graphs and tables, the paper correlates various aspects including education, socio-economic factors, and finance to explain the declining birth rates.
 
-* How does the birth rates vary by the race and ethnicity of mothers'?
+We replicate the study by Kearney, Levine, and Pardue (2022) with a focus on the following research questions:
+
+* What is the trend in US birth rates over the period 1980 to 2020?
+
+* How do birth rates vary by age group?
+
+* How do birth rates vary by the race and ethnicity of mothers?
 

### 2024-02-11 outputs/paper/paper.qmd
Update paper.qmd

@@
 date-format: long
-abstract: "Decreasing birth rates became the conversation topic globally in recent years.
-Kearney, Levine, and Pardue (2022) analyzed possible factors in demographical, economical and social aspects from the period 1980 to 2020.
-By reproducing some applications of the 'The Puzzle of Falling US Birth Rates Since the Great Recession,' this paper demonstrate potential factors correlating aspects on economics, religions, laws and enforcement applied in each states."
+abstract: "This paper investigates the significant decline in US birth rates between 2007 and 2020, replicating the study by Kearney, Levine, and Pardue (2022).
+Analyzing trends from 1980 to 2020, we explore variations in birth rates across age groups and disparities based on the race and ethnicity of mothers.
+While attributing the initial decline to the Great Recession, our analysis identifies unclear factors beyond this period.
+Furthermore, our exploration delves into socio-cultural aspects, providing an understanding of declining birth rates and contributing insights for policymakers and researchers concerned with population trends."
 format: pdf
@@
 ## Methodology
-The data for the trend in US Birth Rates, births per 1,000 women age 15-44 from the period 1980 to 2020 was acquired from CDC Vital Statistics Births Reports for 2015, 2019 and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
-Another trends in Birth Rates by Population Subgroup, specifically birth rates by age group, race and ethnicity are again gathered from CDC Vital Statistics Births Reports.
-Original paper uses single-age population counts, among all races from 1969-2019 and by race and Hispanic origin from 1990-2019, from the CDC SEER database (CDC NCI 2021).
-Data is available freely at https://seer.cancer.gov/popdata/download.html and raw files can be found in the data/pop/folder of the original replication package.
+
+The data concerning the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from the period 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
+Additionally, trends in Birth Rates by Population Subgroup, specifically birth rates categorized by age group, race, and ethnicity, were again gathered from CDC Vital Statistics Births Reports.
+The original paper utilized single-age population counts across all races from 1969 to 2019 and by race and Hispanic origin from 1990 to 2019, sourced from the CDC SEER database (CDC NCI 2021).
+The data is freely available at https://seer.cancer.gov/popdata/download.html, with raw files located in the data/pop/ folder of the original replication package.
 

### 2024-02-11 outputs/paper/paper.qmd
Update paper.qmd

@@
 ## Features
-From now on, every white refers to non-Hispanic, caucassian. Every black refers to non-Hispanic, African American. Hispanic refers to people with Hispanic ethinicity majority from south America.
+De facto population of women of reproductive age is 15 to 49 years in a country (WHO 2024).
+The average age of the final menstrual period (menopause) is age 51, women age in 40s, pregnancy is rare but not impossible (asrm 2012).
+Beginning in 1997, the birth rate for the maternal age group 45-49 includes data for mothers aged 45 and over in the numerator and is based on the population of women aged 45-49 in the denominator (National Center for Health Statistics 2023).
+Thus the women population in the age group 45-49 is eliminated in this paper although they fall under the reproductive age group category in a country.
+Birth data is comprised of births per 1,000 women age 15 to 44 over the 40 years period from 1980 to 2020.
+For the race and ethnicity, every white refers to non-Hispanic, caucassian population group.
+Every black refers to non-Hispanic, African American group.
+Hispanic refers to people with Hispanic ethinicity majority from south America to abberiviate.
 
 # Results
-The results section should convey findings.
-## Table, graph, table, graph
 
-## Source
+## Trend in US Birth Rates
+The overall trend fluctuates from time to time but the highest births moment is recorded as in 1991 for the rate of 71
+
+## Trends in Birth Rates by Population Subgroup
+
 ```{r}
@@
 
-Also, we should not disregard the fact that for all statistical data, they always going to be a missing value that is not properly collected.
-A lot of people might not just report or illegally keep the population due to their illegal status in the US.
-
 Therefore, the data might be skewed due to aforementioned reasons.

### 2024-02-11 outputs/paper/paper.qmd
add scatter plot

--- 2024-02-11 outputs/paper/paper.qmd
@@
   theme_classic()
+```
+
+```{r}
+# Load necessary libraries
+library(ggplot2)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
+
+# Plotting the data with scatter plot and trend line
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_point(color = "blue") +
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2009, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
+  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+
 ```

### 2024-02-11 outputs/paper/paper.qmd
scatter plot
update

--- 2024-02-11 outputs/paper/paper.qmd
@@
 ## Trend in US Birth Rates
-The overall trend fluctuates from time to time but the highest births moment is recorded as in 1991 for the rate of 71
-
-## Trends in Birth Rates by Population Subgroup
+The overall trend fluctuates over time, with the highest number of births recorded in 1990 at a rate of 70.9.
+The second-highest peaks occurred in 1991 and 2007, with rates of 69.3.
+These peaks are accompanied by valleys.
+However, the difference between the birth rates in 1980 and 2020 is 12.6, and the gap between the maximum and minimum points during this period is 15.1.
+These calculations clearly demonstrate the variation over time and suggest that there may be factors influencing the gradual negative slope of the trend.
+Figure 2 presents the scatter plot version of Figure 1, providing a clearer representation of the negative trend observed in the line graph.
 
@@
 
-Figure 1 Trend in US Birth Rates, not just the Great Recession was there another events? 
-abortion law, law enforcement dates, economic crisis other than great recession
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-scatter-graph
+#| fig-cap: "Negative Trend in US Birth Rates"
+
+# Load necessary libraries
+library(ggplot2)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Read the data from CSV file
+data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
+
+# Plotting the data with scatter plot and trend line
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_point(color = "skyblue") +
+  geom_smooth(method = "lm", se = FALSE, color = "pink") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2009, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
+  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+```
+
+## Trends in Birth Rates by Population Subgroup
 
@@
 
-In the discussion section, and any other relevant section, please be sure to discuss ethics and bias, with reference to relevant literature.
+Figure 1 Trend in US Birth Rates, not just the Great Recession was there another events? 
+abortion law, law enforcement dates, economic crisis other than great recession
 

### 2024-02-11 outputs/paper/paper.qmd
update

@@
 Figure 2 presents the scatter plot version of Figure 1, providing a clearer representation of the negative trend observed in the line graph.
+Lastly, the year 2007 has been marked in dotted line to signify the beginning of the fall on both graphs.
 
@@
 ## Trends in Birth Rates by Population Subgroup
+Trends in Birth Rates by Population Subgroup has two panels specified into two: Group A is designated to five-year age group and Group B is designed to illustrate race and ethnicity (ages 15-44).
+
+Both plots present the year on the x-axis and births per 1,000 women in relevant population subgroup in the y-axis.
+For group A, we have 6 different lines to grop the five-year age group of women from the age 15 to 44.
+Again, the Great Recession is marked as a dotted vertical line on both panels.
+While it is generally true that the birth rates drop significantly as of 2007 for the women population group at age between 15 to 29, the elder group at age over 30 shows ascending trends over the year.
 

### 2024-02-11 outputs/paper/paper.qmd
Update some codes

@@
 ```
-Figure 2 A, abortion law applicable age ~,
 
-Technology enhancement in medical field allows the abortion to be held more easily compared to 1980, 1990s.
-This change allow the young women who can afford the abortion or vice versa who can't afford giving birth to get the abortion.
-In the meantime, due to the advancement of technology, women at relatively old age can give the birth in their late 30s or even 40s.
-As well as change in child birth policy and health maintenance was eased compared to the old days in 1980s, 1990s, that promotes women at age older than 30 who are relatively financially stable have high willingness to give births, showing ascending pattern in birth rates over the 40 years.
+# Discussion
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-#| label: tbl-trendssub-table
-#| tbl-cap: "Trends in Birth Rates by Population Race and Ethnicity"
+## Findings
 
-# Load necessary libraries
-library(dplyr)
-library(knitr)
-library(kableExtra)
+In panel A of @fig-bratepop-graph can be interpreted as the proportion of US women using a contraceptive method rose from 56% in 1982 to 60% in 1988 and 64% in 1995.
+This trends in contraceptive use have implications for downward shifts in preganancy rates which ultimately affects birth rates (Piccinino and Mosher 1998).
+Moreover, states began to pass parental involvement laws to limit minors' confidential access to abortion (Myers 2022).
+Meaning that although by 1980 judicial precedent established that minors can consent to abortion services absent a valid restriction, as of 1980, minors had a limitation on abortion which possibly draw correlation between the more usage in contraceptive measurements to prevent the pregnancy.
 
-# Read the data from CSV file
-data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))
+As a result, legal access to abortion and contraceptive in the 1960s and
 
-# Filter the data to include years starting from 1990
-filtered_data <- data %>%
-  filter(year >= 1990)
-
-# Select the required columns and rename them
-selected_data <- filtered_data %>%
-  select(Year = year, White = brate_whitenh, Black = brate_blacknh, Hispanic = brate_hisp)
-
-# Print the table using additional styling
-kable(selected_data, caption = "Trends in Birth Rates by Population Race and Ethnicity (ages 15-44)", format = "html") %>%
-  kable_styling(full_width = FALSE) %>%
-  add_header_above(c("","Births per 1,000 women" = 3))
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-#| label: fig-trendssub-graph
-#| fig-cap: "Trends in Birth Rates by Population Race and Ethnicity"
-
-# Load required libraries
-library(dplyr)
-library(ggplot2)
-
-# Refer to the table a
-# Assume selected_data contains the relevant data
-
-# Create bar graphs for each racial/ethnic group
-bar_plot <- ggplot(selected_data, aes(x = Year)) +
-  geom_bar(aes(y = Hispanic, fill = "Hispanic"), stat = "identity", position = "stack") +
-  geom_bar(aes(y = Black, fill = "Black"), stat = "identity", position = "stack") +
-  geom_bar(aes(y = White, fill = "White"), stat = "identity", position = "stack") +
-  labs(x = "Year",
-       y = "Births per 1,000 women relevant population subgroup") +
-  scale_fill_manual(values = c(Hispanic = "grey", Black = "orange", White = "blue"),
-                    name = "Race and ethnicity") +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(1990, 2020, by = 5), minor_breaks = seq(1990, 2020, by = 1)) +
-  ylim(0, 120)
-
-# Print the bar plot
-print(bar_plot)
-```
-Figure 3 B, major religion of Hispanic, Black, White  religious reason of each race and ethnicity.
+In panel B of @fig-bratepop-graph, major religion of Hispanic, Black, White  religious reason of each race and ethnicity.
 
@@
 -whether they accept the abortion or not, cultural perspectives, whether marriage affects the decision of keeping a child or not.<- financial reason
-
-Might include Map or not depending on the sources and references I find, see how long it'll gonna be.
-
-# Discussion
-
-Figure 1 Trend in US Birth Rates, not just the Great Recession was there another events? 
-abortion law, law enforcement dates, economic crisis other than great recession
-
-## Findings
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 

### 2024-02-11 outputs/paper/paper.qmd
update

@@
 Again, the Great Recession is marked as a dotted vertical line on both panels.
-While it is generally true that the birth rates drop significantly as of 2007 for the women population group at age between 15 to 29, the elder group at age over 30 shows ascending trends over the year.
+While it is generally true that the birth rates drop significantly as of 2007 for the women population group at age between 15 to 29, the elder group at age over 30 shows ascending trends over the year regardless of the event.
+Also, the age group 15-29 seems to be following downward slope from 1980 until it hits its maximum in 1991.
+
+In panel B of @fig-bratepop-graph has three different subgroups. Race and ethnicity is subdivided into three sections, Hispanic, Black, and White. Both white and black are non-Hispanic population.
 
@@
 
+According to population and development review,
+
 As a result, legal access to abortion and contraceptive in the 1960s and

### 2024-02-12 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-In panel B of @fig-bratepop-graph has three different subgroups. Race and ethnicity is subdivided into three sections, Hispanic, Black, and White. Both white and black are non-Hispanic population.
+Note that the data set is gathered as of 1989 for the race and ethnicity subgroup population, thus the x-axis of the graph starts from 1989.
+In panel B of @fig-bratepop-graph has three different subgroups.
+Race and ethnicity is subdivided into three sections, Hispanic, Black, and White.
+Both white and black are non-Hispanic population.
+The white population group shows relatively linear fashion compared to other population groups while the birth rates of Hispanic and Black population groups drastically drop over the 10 years from 1989 to 2000.
+Then another significant drop happened as of 2007 only in Hispanic population group.
+The change in births per 1,000 women in race and ethnicity population subgroup varies, 7.3, 25.8 and 42.1 in White, Black, and Hispanic group respectively.
 

### 2024-02-12 outputs/paper/paper.qmd
Update paper.qmd
Update

@@
 Both plots present the year on the x-axis and births per 1,000 women in relevant population subgroup in the y-axis.
-For group A, we have 6 different lines to grop the five-year age group of women from the age 15 to 44.
+For group A, we have 6 different lines to group the five-year age group of women from the age 15 to 44.
 Again, the Great Recession is marked as a dotted vertical line on both panels.
@@
 Both white and black are non-Hispanic population.
-The white population group shows relatively linear fashion compared to other population groups while the birth rates of Hispanic and Black population groups drastically drop over the 10 years from 1989 to 2000.
+While the white population group shows relatively linear fashion over 30 years, Hispanic population birth groups were 1.7 times and black population birth groups being 1.3 times larger than White population group up until 2007.
+The birth rates of Hispanic and Black population groups drastically drop over the 10 years from 1989 to 2000.
 Then another significant drop happened as of 2007 only in Hispanic population group.
 The change in births per 1,000 women in race and ethnicity population subgroup varies, 7.3, 25.8 and 42.1 in White, Black, and Hispanic group respectively.
+As of 2012, all race and ethnicity population group illustrate the relatively stable tendency and slowly declining.
 
@@
 
-According to population and development review,
+In panel B of @fig-bratepop-graph, Hispanic group
 
-As a result, legal access to abortion and contraceptive in the 1960s and
+Regarding abortion, contraceptive use, and sexual activity before marriage:
 
-In panel B of @fig-bratepop-graph, major religion of Hispanic, Black, White  religious reason of each race and ethnicity.
+ **White Evangelical Protestants**:
+   - Generally oppose abortion and view it as morally wrong, often advocating for anti-abortion legislation.
+   - Emphasize abstinence before marriage and may discourage or oppose contraceptive use outside of marriage.
 
-Here we are dissecting the cultural and religious difference between the races among the population subgroup. 
-Hispanic, majority catholic,
-White, christian, catholic, etc
-Black, christian, catholic, etc
+ **White Mainline (Non-Evangelical) Protestants**:
+   - Views on abortion and contraceptive use may vary within this group, but tend to be less dogmatic compared to evangelical Protestants.
+   - May hold a range of views on sexual activity before marriage, with some advocating for abstinence and others accepting of premarital sexual relationships.
 
-It is prevailed that certain religion does not allow the abortion, saying that it disobey the words of God or whatever.
+ **Black Protestants**:
+   - Views on abortion and contraceptive use may vary within this group, but many Black Protestant denominations are socially conservative and oppose abortion.
+   - Teachings on sexual activity before marriage may emphasize abstinence, although individual beliefs and practices may vary.
 
--whether they accept the abortion or not, cultural perspectives, whether marriage affects the decision of keeping a child or not.<- financial reason
+ **Hispanic Catholics**:
+   -  Official Catholic doctrine opposes abortion and teaches that life begins at conception, making abortion morally unacceptable.
+   - Contraceptive use is a topic of debate within the Catholic Church, with the official stance opposing artificial contraception, although many Catholics use contraceptives despite this teaching.
+   - Catholic teachings emphasize abstinence before marriage but recognize that many Catholics engage in premarital sexual activity.
 
 ## Ethical Implication
+It's important to note that beliefs and practices related to abortion, contraceptive use, and sexual activity before marriage can vary within and among different religious groups.
+Additionally, individual beliefs and practices may not always align perfectly with the official teachings or doctrines of a particular religious tradition.
 

### 2024-02-12 outputs/paper/paper.qmd
update

@@
 
-In panel B of @fig-bratepop-graph, Hispanic group
+According to the article from the journal of the National Medical Association, many Hispanic people in the United States retain the customs and traditions of their country of origin.
+As a result, their use of preventive health care measures is lower than the majority of the population.
+Moreover, Hispanics often marry and start a family earlier than the average US citizen which corresponds higher birth rate and a higer fertility rate (number of live births per 1,000 women aged 15 to 44 years) than the average.
+Moreover, Hispanic culture expect women to bear children.
+Abortion, adoption, and artificial insemination are less acceptable to Hispanic than to the average American (Poma 1987).
+That explains why Hispanic births rate were much higher comopared to the other population groups in panel B of @fig-bratepop-graph.
 
-Regarding abortion, contraceptive use, and sexual activity before marriage:
-
- **White Evangelical Protestants**:
-   - Generally oppose abortion and view it as morally wrong, often advocating for anti-abortion legislation.
-   - Emphasize abstinence before marriage and may discourage or oppose contraceptive use outside of marriage.
-
- **White Mainline (Non-Evangelical) Protestants**:
-   - Views on abortion and contraceptive use may vary within this group, but tend to be less dogmatic compared to evangelical Protestants.
-   - May hold a range of views on sexual activity before marriage, with some advocating for abstinence and others accepting of premarital sexual relationships.
-
- **Black Protestants**:
-   - Views on abortion and contraceptive use may vary within this group, but many Black Protestant denominations are socially conservative and oppose abortion.
-   - Teachings on sexual activity before marriage may emphasize abstinence, although individual beliefs and practices may vary.
-
- **Hispanic Catholics**:
-   -  Official Catholic doctrine opposes abortion and teaches that life begins at conception, making abortion morally unacceptable.
-   - Contraceptive use is a topic of debate within the Catholic Church, with the official stance opposing artificial contraception, although many Catholics use contraceptives despite this teaching.
-   - Catholic teachings emphasize abstinence before marriage but recognize that many Catholics engage in premarital sexual activity.
+Christianity, including protestants and catholic, generally oppose the idea of abortion and view it as morally wrong.
+They also oppose contraceptive use outside of marriage.
+Religious group are socially conservative.
+With this religious reason behind it, the social norm was opposed to the idea of abortion, contraceptive use back in the days which is why this can be a reason why the birth rate is decreasing for all race and ethnicity groups.
+Shift in religious activities of young Americans.
+Early 1990s, about 90% of US adults identified as Christians.
+However, over the time, young Americans are now less likely to become or remain Christian (Pew Research Center 2022).
+In 2020, unaffiliated group share the greatest part of the American religious landscape in 2020 by parting 23% (PRRI 2020).
 
@@
 ## Future Research
-It might be interesting to expand the research worldwide, not just limiting ourselves into the US data.
-It is broadly known that Chinese and Indian population are growing exponentially while the birth rate in other developed countries continue to grow small.
-It might be interesting to research more what's behind.
+Political demographic, immigration law, and enforcement.
 

### 2024-02-12 outputs/paper/paper.qmd
update

@@
 Many countries are addressing this issue of declining birth rates and proposing alternative measures to compensate for the population drop.
-For instance, Germany's birth rate (births per 1000 people) in 2024 is 9.3, which is lower than Canada's 10.0 and the US's 12.0 (Macrotrends 2024).
+For instance, Germany's birth rate in 2024 is 9.3, which is lower than Canada's 10.0 and the US's 12.0 (Macrotrends 2024).
 Germany is taking action by actively recruiting skilled workers from abroad (Grunau 2023) and reforming its citizenship law to allow dual citizenship (Knight 2024) as some measures to combat declining birth rates and overall population.
@@
 This paper utilized a replication package of the original study and employed a cleaned data CSV file to reproduce the related figures and tables.
-This approach enabled us to draw correlations between the dataset and political, economic, and social policy measures that were not covered in the original paper.
-Furthermore, this paper investigates various socio-cultural aspects, including sexual education, marriage rates, financial stability among women over the age of 30, and the role of technology in healthcare.
-
--what was found-
+This approach enabled us to draw correlations between the dataset and
 
@@
 Beginning in 1997, the birth rate for the maternal age group 45-49 includes data for mothers aged 45 and over in the numerator and is based on the population of women aged 45-49 in the denominator (National Center for Health Statistics 2023).
-Thus the women population in the age group 45-49 is eliminated in this paper although they fall under the reproductive age group category in a country.
+Thus the women population in the age group 45-49 is eliminated in this paper although they fall under the reproductive age group category.
 Birth data is comprised of births per 1,000 women age 15 to 44 over the 40 years period from 1980 to 2020.
@@
 These calculations clearly demonstrate the variation over time and suggest that there may be factors influencing the gradual negative slope of the trend.
-Figure 2 presents the scatter plot version of Figure 1, providing a clearer representation of the negative trend observed in the line graph.
+@fig-scatter-graph presents the scatter plot version of @fig-bratetrend-graph, providing a clearer representation of the negative trend observed in the line graph.
 Lastly, the year 2007 has been marked in dotted line to signify the beginning of the fall on both graphs.
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-polscatter-table
+#| tbl-cap: "US presidents"
+
+# Load necessary libraries
+library(ggplot2)
+
+# Create a dataframe with information about the Presidents and their respective parties
+presidents <- data.frame(
+  president = c("Ronald Reagan", "George H. W. Bush", "Bill Clinton", "George W. Bush", "Barack Obama", "Donald Trump", "Joe Biden"),
+  party = c("Republican", "Republican", "Democratic", "Republican", "Democratic", "Republican", "Democratic"),
+  start_year = c(1981, 1989, 1993, 2001, 2009, 2017, 2021),
+  end_year = c(1989, 1993, 2001, 2009, 2017, 2021, NA) # NA for current President
+)
+
+head(presidents)
+
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-polscatter-graph
+#| fig-cap: "Trend in US Birth Rates with political party in their presidential mandate"
+
+# Load necessary libraries
+library(ggplot2)
+
+# Plotting the data with scatter plot and trend lines for each presidential mandate period
+ggplot(data, aes(x = year, y = brate_all)) +
+  geom_point(color = "pink") +
+  geom_smooth(data = subset(data, year >= 1980 & year <= 1981), method = "lm", se = FALSE, color = "blue") +
+  geom_smooth(data = subset(data, year >= 1981 & year <= 1989), method = "lm", se = FALSE, color = "red") +
+  geom_smooth(data = subset(data, year >= 1989 & year <= 1993), method = "lm", se = FALSE, color = "red") +
+  geom_smooth(data = subset(data, year >= 1993 & year <= 2001), method = "lm", se = FALSE, color = "blue") +
+  geom_smooth(data = subset(data, year >= 2001 & year <= 2009), method = "lm", se = FALSE, color = "red") +
+  geom_smooth(data = subset(data, year >= 2009 & year <= 2017), method = "lm", se = FALSE, color = "blue") +
+  geom_smooth(data = subset(data, year >= 2017 & year <= 2021), method = "lm", se = FALSE, color = "red") +
+  geom_vline(xintercept = 2007, linetype = "dotted") +
+  annotate("text", x = 2009, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
+  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
+  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
+  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
+  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
+  theme_classic()
+
+```
+
 ## Trends in Birth Rates by Population Subgroup
-Trends in Birth Rates by Population Subgroup has two panels specified into two: Group A is designated to five-year age group and Group B is designed to illustrate race and ethnicity (ages 15-44).
+@fig-bratepop-graph has two panels specified into two: Group A is designated to five-year age group and Group B is designed to illustrate race and ethnicity (ages 15-44).
 
@@
 While the white population group shows relatively linear fashion over 30 years, Hispanic population birth groups were 1.7 times and black population birth groups being 1.3 times larger than White population group up until 2007.
-The birth rates of Hispanic and Black population groups drastically drop over the 10 years from 1989 to 2000.
+The birth rates of Hispanic and Black population groups drastically drop over the 11 years from 1989 to 2000.
 Then another significant drop happened as of 2007 only in Hispanic population group.
@@
 
-Christianity, including protestants and catholic, generally oppose the idea of abortion and view it as morally wrong.
+Another interesting factor could be christianity.
+Including protestants and catholic, christian generally against with the idea of abortion and view it as morally wrong.
 They also oppose contraceptive use outside of marriage.
 Religious group are socially conservative.
-With this religious reason behind it, the social norm was opposed to the idea of abortion, contraceptive use back in the days which is why this can be a reason why the birth rate is decreasing for all race and ethnicity groups.
-Shift in religious activities of young Americans.
+However, there was a shift in religious activities of young Americans.
 Early 1990s, about 90% of US adults identified as Christians.
@@
 In 2020, unaffiliated group share the greatest part of the American religious landscape in 2020 by parting 23% (PRRI 2020).
+This change in religious activity could explain the reason why the birth rate is decreasing for all race and ethnicity groups over the period.
 
@@
 ## Future Research
-Political demographic, immigration law, and enforcement.
+It might be worth investigating some other aspects of the society such as political demographic, immigration law, and enforcement. Based on trend in
 

### 2024-02-12 outputs/paper/paper.qmd
update

@@
 #| message: false
-#| label: tbl-polscatter-table
-#| tbl-cap: "US presidents"
-
-# Load necessary libraries
-library(ggplot2)
-
-# Create a dataframe with information about the Presidents and their respective parties
-presidents <- data.frame(
-  president = c("Ronald Reagan", "George H. W. Bush", "Bill Clinton", "George W. Bush", "Barack Obama", "Donald Trump", "Joe Biden"),
-  party = c("Republican", "Republican", "Democratic", "Republican", "Democratic", "Republican", "Democratic"),
-  start_year = c(1981, 1989, 1993, 2001, 2009, 2017, 2021),
-  end_year = c(1989, 1993, 2001, 2009, 2017, 2021, NA) # NA for current President
-)
-
-head(presidents)
-
-```
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
 #| label: fig-polscatter-graph
@@
 
+@fig-polscatter-graph draw multiple trend lines
+
 In panel A of @fig-bratepop-graph can be interpreted as the proportion of US women using a contraceptive method rose from 56% in 1982 to 60% in 1988 and 64% in 1995.
@@
 ## Future Research
-It might be worth investigating some other aspects of the society such as political demographic, immigration law, and enforcement. Based on trend in
+It might be worth investigating some other aspects of the society such as political demographic, immigration law, and enforcement.
+Based on trend in @fig-polscatter-graph, it is worth to delve into immigration law, campaign promises for their presidential mandates.
 

### 2024-02-12 outputs/paper/paper.qmd
update

@@
 ggplot(data, aes(x = year, y = brate_all)) +
-  geom_point(color = "skyblue") +
-  geom_smooth(method = "lm", se = FALSE, color = "pink") +
+  geom_point(color = "pink") +
+  geom_smooth(method = "lm", se = FALSE, color = "skyblue") +
   geom_vline(xintercept = 2007, linetype = "dotted") +

### 2024-02-12 outputs/paper/paper.qmd
update

@@
 This paper utilized a replication package of the original study and employed a cleaned data CSV file to reproduce the related figures and tables.
-This approach enabled us to draw correlations between the dataset and
+This approach enabled us to draw correlations between the dataset and socio-cultural asepcts including abortion, contraceptive use, financial stability, religion, culture.
 
@@
 #| label: fig-polscatter-graph
-#| fig-cap: "Trend in US Birth Rates with political party in their presidential mandate"
+#| fig-cap: "Trend in US Birth Rates during their presidential mandate"
 
@@
 
-@fig-polscatter-graph draw multiple trend lines
+@fig-polscatter-graph draw multiple trend lines in US birth rates during their presidential mandate.
+It also indicates their political party, blue represents the democrat while red presents the republican.
+Radican drop while republican period is seen from the analysis while democrat presidential mandate period illustrates relatively moderate delicining tendency.
 
@@
 
-## Accounting for bias
-
 ## Limitation

### 2024-02-12 outputs/paper/paper.qmd
editing words

@@
 ## Methodology
-
-The data concerning the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from the period 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
-Additionally, trends in Birth Rates by Population Subgroup, specifically birth rates categorized by age group, race, and ethnicity, were again gathered from CDC Vital Statistics Births Reports.
+The data on the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
+Additionally, trends in Birth Rates by Population Subgroup, specifically birth rates categorized by age group, race, and ethnicity, were gathered from the CDC Vital Statistics Births Reports.
 The original paper utilized single-age population counts across all races from 1969 to 2019 and by race and Hispanic origin from 1990 to 2019, sourced from the CDC SEER database (CDC NCI 2021).
@@
 It also indicates their political party, blue represents the democrat while red presents the republican.
-Radican drop while republican period is seen from the analysis while democrat presidential mandate period illustrates relatively moderate delicining tendency.
+Radical drop while republican period is seen from the analysis while democrat presidential mandate period illustrates relatively moderate delicining tendency.
 
@@
 Meaning that although by 1980 judicial precedent established that minors can consent to abortion services absent a valid restriction, as of 1980, minors had a limitation on abortion which possibly draw correlation between the more usage in contraceptive measurements to prevent the pregnancy.
+The reason behind the ascending patterns of the age group over 30 can be explained due to advancement of health care and financial stability.
 

### 2024-02-12 outputs/paper/paper.qmd
update

--- 2024-02-12 outputs/paper/paper.qmd
@@
 date-format: long
-abstract: "This paper investigates the significant decline in US birth rates between 2007 and 2020, replicating the study by Kearney, Levine, and Pardue (2022).
-Analyzing trends from 1980 to 2020, we explore variations in birth rates across age groups and disparities based on the race and ethnicity of mothers.
-While attributing the initial decline to the Great Recession, our analysis identifies unclear factors beyond this period.
-Furthermore, our exploration delves into socio-cultural aspects, providing an understanding of declining birth rates and contributing insights for policymakers and researchers concerned with population trends."
+abstract: "This paper examines the notable decrease in US birth rates from 2007 to 2020, replicating the study conducted by Kearney, Levine, and Pardue (2022).
+Through an analysis of trends spanning from 1980 to 2020, we investigate variations in birth rates among different age groups and disparities based on the race and ethnicity of mothers.
+While initially attributing the decline to the Great Recession, our analysis uncovers additional factors contributing to this trend.
+Furthermore, our exploration delves into socio-cultural aspects, offering insights for policymakers and researchers interested in population trends and their implications."
 format: pdf
@@
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
-```
-
 # Introduction
@@
 From an economic perspective, birth rates are an important factor that determines the growth of a country.
-In recent years, the low birth rate has become a conversation topic both nationwide and worldwide.
-Many countries are addressing this issue of declining birth rates and proposing alternative measures to compensate for the population drop.
+Recent years have seen widespread concern over declining birth rates globally, prompting nations to explore strategies to address this issue.
 For instance, Germany's birth rate in 2024 is 9.3, which is lower than Canada's 10.0 and the US's 12.0 (Macrotrends 2024).
 Germany is taking action by actively recruiting skilled workers from abroad (Grunau 2023) and reforming its citizenship law to allow dual citizenship (Knight 2024) as some measures to combat declining birth rates and overall population.
-Finally, the birth rate also serves as an indicator of replacement level fertility, which in developed countries can be taken as requiring an average of 2.1 children per woman (Craig 1994).
-Therefore, birth rates are poised to become a key global topic in the mid to long term.
+Furthermore, birth rates serve as an indicator of replacement-level fertility, essential for maintaining population stability in developed nations (Craig 1994).
 
@@
 While they attributed the main cause of the decline in the early part of the period to the Great Recession, they were unable to identify any other economic, policy, or social factors that have changed since 2007 to support the decline beyond the Great Recession (Kearney, Levine, and Pardue 2022).
-The objective of this paper is to replicate the figures and tables from the original study and to explore additional factors contributing to the declining birth rates in the US.
-In addition to examining the impact of the Great Recession, this paper delves into the perspectives of population subgroups, analyzing data in five-year age groups and by race and ethnicity to gain deeper insights.
+This paper aims to replicate and expand upon their study by examining additional factors influencing birth rate trends in the US.
+In particular, it explores the perspectives of various population subgroups, analyzing data by age groups and race/ethnicity to gain deeper insights.
 
 This paper utilized a replication package of the original study and employed a cleaned data CSV file to reproduce the related figures and tables.
-This approach enabled us to draw correlations between the dataset and
+While the original code in the replication package was written in Stata, this paper utilized R (R Core Team 2020) for analysis, leveraging packages such as ggplot2 (Wickham et al.
+2016) for creating graphical representations.
+Additionally, the gridExtra package was used to enhance the layout and presentation of the figures (Auguie 2017).
 
-First, the paper showcases the trend in US birth rates and discusses the relevant factors along with the figure.
-It then demonstrates trends in birth rates by population subgroup, focusing on two key areas: five-year age groups and race and ethnicity.
-Alongside the graphs and tables, the paper correlates various aspects including education, socio-economic factors, and finance to explain the declining birth rates.
+The paper proceeds by first presenting the trend in US birth rates over the period 1980 to 2020, accompanied by relevant factors (CDC).
+It then delves into variations in birth rates by age group and race/ethnicity of mothers, correlating these trends with education, socio-economic factors, and financial stability.
+The replication of Kearney, Levine, and Pardue's study focuses on answering the following research questions:
 
-We replicate the study by Kearney, Levine, and Pardue (2022) with a focus on the following research questions:
+* What is the trend in US birth rates from 1980 to 2020?
 
-* What is the trend in US birth rates over the period 1980 to 2020?
+* How do birth rates vary across different age groups?
 
-* How do birth rates vary by age group?
+* How do birth rates differ based on the race and ethnicity of mothers?
 
-* How do birth rates vary by the race and ethnicity of mothers?
-
-The original code in the replication package is written in Stata.
-However, this paper utilizes R (R Core Team 2020) to analyze a dataset from the replication package.
-Packages such as ggplot2 (Wickham et al.
-2016) for creating graphical representations, kableExtra (Zhu et al.
-2021) for enhancing table aesthetics, and lubridate (Grolemund et al.
-2021) for handling date-related operations were employed.
-Additionally, the Tidyverse framework (Wickham et al.
-2019) ensures cohesiveness, while Dplyr (Wickham et al.
-2021) facilitates data manipulation.
+Through this comprehensive analysis, we aim to contribute to a better understanding of the complex factors influencing birth rate trends in the United States.
 
@@
 
-The data concerning the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from the period 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
-Additionally, trends in Birth Rates by Population Subgroup, specifically birth rates categorized by age group, race, and ethnicity, were again gathered from CDC Vital Statistics Births Reports.
+The data on the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
+Additionally, trends in Birth Rates by Population Subgroup, specifically birth rates categorized by age group, race, and ethnicity, were gathered from the CDC Vital Statistics Births Reports.
 The original paper utilized single-age population counts across all races from 1969 to 2019 and by race and Hispanic origin from 1990 to 2019, sourced from the CDC SEER database (CDC NCI 2021).
@@
 ## Features
-De facto population of women of reproductive age is 15 to 49 years in a country (WHO 2024).
-The average age of the final menstrual period (menopause) is age 51, women age in 40s, pregnancy is rare but not impossible (asrm 2012).
+
+The de facto population of women of reproductive age spans from 15 to 49 years in a country (WHO 2024).
+The average age of the final menstrual period (menopause) is 51, and for women in their 40s, pregnancy is rare but not impossible (ASRM 2012).
 Beginning in 1997, the birth rate for the maternal age group 45-49 includes data for mothers aged 45 and over in the numerator and is based on the population of women aged 45-49 in the denominator (National Center for Health Statistics 2023).
-Thus the women population in the age group 45-49 is eliminated in this paper although they fall under the reproductive age group category.
-Birth data is comprised of births per 1,000 women age 15 to 44 over the 40 years period from 1980 to 2020.
-For the race and ethnicity, every white refers to non-Hispanic, caucassian population group.
-Every black refers to non-Hispanic, African American group.
-Hispanic refers to people with Hispanic ethinicity majority from south America to abberiviate.
+Thus, the women population in the age group 45-49 is excluded in this paper, although they fall under the reproductive age group category.
+Birth data comprises births per 1,000 women aged 15 to 44 over the 40-year period from 1980 to 2020.
+
+Regarding race and ethnicity, "white" refers to the non-Hispanic Caucasian population group, "black" refers to the non-Hispanic African American group, and "Hispanic" refers to people with Hispanic ethnicity, mainly from South America, for brevity.
 
@@
 ## Trend in US Birth Rates
-The overall trend fluctuates over time, with the highest number of births recorded in 1990 at a rate of 70.9.
+
+The overall trend in US Birth Rates fluctuates over time, with the highest number of births recorded in 1990 at a rate of 70.9.
 The second-highest peaks occurred in 1991 and 2007, with rates of 69.3.
@@
 @fig-scatter-graph presents the scatter plot version of @fig-bratetrend-graph, providing a clearer representation of the negative trend observed in the line graph.
-Lastly, the year 2007 has been marked in dotted line to signify the beginning of the fall on both graphs.
+Lastly, the year 2007 has been marked with a dotted line to signify the beginning of the fall on both graphs.
 
@@
   scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
-  labs(x = "Year", y = "Births per 1,000 women age 15-44",
+  labs(x = NULL, y = "Births per 1,000 women age 15-44",
        caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
-  theme_classic()
+  theme_classic() +
+  theme(text = element_text(family = "Times New Roman"))
+
 ```
@@
 #| label: fig-polscatter-graph
-#| fig-cap: "Trend in US Birth Rates with political party in their presidential mandate"
+#| fig-cap: "Trend in US Birth Rates during the presidential mandate"
+
+# Set seed for reproducibility
+set.seed(123)
 
@@
   scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
-  labs(x = "Year", y = "Births per 1,000 women age 15-44",
-       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
-  theme_classic()
+  labs(x = NULL, y = "Births per 1,000 women age 15-44",
+       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.", 
+       title = NULL) +
+  theme_classic() +
+  theme(text = element_text(family = "Times New Roman"))
 
@@
 ## Trends in Birth Rates by Population Subgroup
-@fig-bratepop-graph has two panels specified into two: Group A is designated to five-year age group and Group B is designed to illustrate race and ethnicity (ages 15-44).
 
-Both plots present the year on the x-axis and births per 1,000 women in relevant population subgroup in the y-axis.
-For group A, we have 6 different lines to group the five-year age group of women from the age 15 to 44.
-Again, the Great Recession is marked as a dotted vertical line on both panels.
-While it is generally true that the birth rates drop significantly as of 2007 for the women population group at age between 15 to 29, the elder group at age over 30 shows ascending trends over the year regardless of the event.
-Also, the age group 15-29 seems to be following downward slope from 1980 until it hits its maximum in 1991.
+@fig-bratepop-graph consists of two panels, each representing distinct aspects of the data.
+Panel Group A is dedicated to depicting birth rates across five-year age groups, while Panel Group B illustrates birth rates by race and ethnicity for women aged 15 to 44.
 
-Note that the data set is gathered as of 1989 for the race and ethnicity subgroup population, thus the x-axis of the graph starts from 1989.
-In panel B of @fig-bratepop-graph has three different subgroups.
-Race and ethnicity is subdivided into three sections, Hispanic, Black, and White.
-Both white and black are non-Hispanic population.
-While the white population group shows relatively linear fashion over 30 years, Hispanic population birth groups were 1.7 times and black population birth groups being 1.3 times larger than White population group up until 2007.
-The birth rates of Hispanic and Black population groups drastically drop over the 11 years from 1989 to 2000.
-Then another significant drop happened as of 2007 only in Hispanic population group.
-The change in births per 1,000 women in race and ethnicity population subgroup varies, 7.3, 25.8 and 42.1 in White, Black, and Hispanic group respectively.
-As of 2012, all race and ethnicity population group illustrate the relatively stable tendency and slowly declining.
+Both plots feature the year on the x-axis and births per 1,000 women in the respective population subgroup on the y-axis.
+In Group A, six different lines delineate the birth rates across five-year age groups from ages 15 to 44.
+The onset of the Great Recession is denoted by a dotted vertical line on both panels.
+Notably, while birth rates significantly decreased after 2007 for women aged 15 to 29, older age groups over 30 exhibited ascending trends regardless of this economic event.
+Additionally, the age group 15-29 showed a downward slope from 1980 until reaching its peak in 1991.
+
+It is important to note that the data set for race and ethnicity subgroups is available from 1989 onwards, hence the x-axis of the graph starts from 1989.
+In Panel B of @fig-bratepop-graph, three distinct subgroups are presented: Hispanic, Black, and White.
+Both the White and Black subgroups represent non-Hispanic populations.
+While the birth rates for the White population group display a relatively linear trend over 30 years, Hispanic and Black population groups exhibited birth rates 1.7 times and 1.3 times higher than the White population group, respectively, until 2007.
+The birth rates of Hispanic and Black population groups experienced significant drops over the 11-year period from 1989 to 2000, followed by another significant decrease as of 2007, observed only in the Hispanic population group.
+The change in births per 1,000 women in race and ethnicity population subgroups varied, with declines of 7.3, 25.8, and 42.1 in the White, Black, and Hispanic groups, respectively.
+As of 2012, all race and ethnicity population groups exhibit a relatively stable trend with a gradual decline.
 
@@
      family = "Times") +
+theme(axis.title.y = element_text(size = 12, family = "Times")) +
   theme_classic() +
@@
 
-@fig-polscatter-graph draw multiple trend lines
+@fig-polscatter-graph displays multiple trend lines representing US birth rates during presidential mandates, with each line corresponding to the political party of the president.
+The trend lines are color-coded, with blue indicating Democrat and red representing Republican presidencies.
+A notable observation from the analysis is a radical drop in birth rates during Republican presidential mandates, whereas Democrat presidential mandate periods illustrate a relatively moderate declining tendency.
 
-In panel A of @fig-bratepop-graph can be interpreted as the proportion of US women using a contraceptive method rose from 56% in 1982 to 60% in 1988 and 64% in 1995.
-This trends in contraceptive use have implications for downward shifts in preganancy rates which ultimately affects birth rates (Piccinino and Mosher 1998).
-Moreover, states began to pass parental involvement laws to limit minors' confidential access to abortion (Myers 2022).
-Meaning that although by 1980 judicial precedent established that minors can consent to abortion services absent a valid restriction, as of 1980, minors had a limitation on abortion which possibly draw correlation between the more usage in contraceptive measurements to prevent the pregnancy.
+In panel A of @fig-bratepop-graph, the trend indicates that the proportion of US women using contraceptive methods increased from 56% in 1982 to 60% in 1988 and further to 64% in 1995.
+These trends in contraceptive use have implications for downward shifts in pregnancy rates, which ultimately affect birth rates (Piccinino and Mosher 1998).
+Additionally, states began passing parental involvement laws to limit minors' confidential access to abortion (Myers 2022).
+This means that although judicial precedent established in 1980 allowed minors to consent to abortion services in the absence of valid restrictions, by 1980 minors faced limitations on abortion, possibly leading to increased usage of contraceptive measures to prevent pregnancy.
+The ascending patterns observed in the age group over 30 can be attributed to advancements in healthcare and financial stability.
 
-According to the article from the journal of the National Medical Association, many Hispanic people in the United States retain the customs and traditions of their country of origin.
-As a result, their use of preventive health care measures is lower than the majority of the population.
-Moreover, Hispanics often marry and start a family earlier than the average US citizen which corresponds higher birth rate and a higer fertility rate (number of live births per 1,000 women aged 15 to 44 years) than the average.
-Moreover, Hispanic culture expect women to bear children.
-Abortion, adoption, and artificial insemination are less acceptable to Hispanic than to the average American (Poma 1987).
-That explains why Hispanic births rate were much higher comopared to the other population groups in panel B of @fig-bratepop-graph.
+According to an article from the Journal of the National Medical Association, many Hispanic people in the United States retain the customs and traditions of their country of origin.
+As a result, their use of preventive healthcare measures is lower than that of the majority of the population.
+Moreover, Hispanics often marry and start a family earlier than the average US citizen, resulting in a higher birth rate and fertility rate (number of live births per 1,000 women aged 15 to 44 years) than the average.
+Additionally, Hispanic culture expects women to bear children, and abortion, adoption, and artificial insemination are less acceptable to Hispanics than to the average American (Poma 1987).
+This explains why Hispanic birth rates were much higher compared to other population groups in panel B of @fig-bratepop-graph.
 
-Another interesting factor could be christianity.
-Including protestants and catholic, christian generally against with the idea of abortion and view it as morally wrong.
-They also oppose contraceptive use outside of marriage.
-Religious group are socially conservative.
-However, there was a shift in religious activities of young Americans.
-Early 1990s, about 90% of US adults identified as Christians.
-However, over the time, young Americans are now less likely to become or remain Christian (Pew Research Center 2022).
-In 2020, unaffiliated group share the greatest part of the American religious landscape in 2020 by parting 23% (PRRI 2020).
+Another interesting factor could be Christianity.
+Including Protestants and Catholics, Christians generally oppose the idea of abortion and view it as morally wrong.
+They also oppose contraceptive use outside of marriage, as religious groups tend to be socially conservative.
+However, there has been a shift in religious activities among young Americans.
+In the early 1990s, about 90% of US adults identified as Christians.
+Over time, however, young Americans are now less likely to become or remain Christian (Pew Research Center, 2022).
+In 2020, the unaffiliated group comprised the largest portion of the American religious landscape, accounting for 23% (PRRI, 2020).
 This change in religious activity could explain the reason why the birth rate is decreasing for all race and ethnicity groups over the period.
+Regardless of religion, people engage in sexual activities before marriage, but as the population with religious affiliations becomes more unaffiliated, they are free to use contraceptives and seek abortion if necessary.
 
 ## Ethical Implication
+
 It's important to note that beliefs and practices related to abortion, contraceptive use, and sexual activity before marriage can vary within and among different religious groups.
@@
 
-## Accounting for bias
+## Limitation
 
-## Limitation
-Beginning in 1997, the birth rate for the maternal age group 45-49 includes data for mothers aged 45 and over in the numerator and is based on the population of women aged 45-49 in the denominator (National Center for Health Statistics 2023).
-However as the dataset is missing for the population of women aged 45-49 before 1997 and the paper counts the data from the period 1980 to 2020, the women population in the age group 45-49 is eliminated.
+Beginning in 1997, the birth rate for the maternal age group 45-49 includes data for mothers aged 45 and over in the numerator and is based on the population of women aged 45-49 in the denominator (National Center for Health Statistics, 2023).
+However, since the dataset is missing data for the population of women aged 45-49 before 1997, and the paper covers the period from 1980 to 2020, the population subgroup of women aged 45-49 is eliminated from the analysis.
 
-As mentioned in the original paper, the population subgroup can be diverged as they mix over the time.
-It is hard to sort out the race completely once they start to mixing it up.
-Therefore, over the time, we don't know the population subgroup Hispanic, Black, White are purely Hispanic, Black, White throughout the period.
+Additionally, the population subgroups can become increasingly mixed over time, making it difficult to accurately classify individuals into distinct racial categories.
+Therefore, it is unclear whether individuals classified as Hispanic, Black, or White remained purely within those categories throughout the period studied.
 
-Therefore, the data might be skewed due to aforementioned reasons.
+As a result of these factors, the data may be skewed, and caution should be exercised when interpreting the results.
 
 ## Future Research
-It might be worth investigating some other aspects of the society such as political demographic, immigration law, and enforcement.
-Based on trend in @fig-polscatter-graph, it is worth to delve into immigration law, campaign promises for their presidential mandates.
+
+For future research, exploring additional aspects of society such as political demographics, immigration law, and enforcement, based on the trends observed in @fig-polscatter-graph, could provide valuable insights.
+Understanding the potential impact of immigration laws and campaign promises during presidential mandates on birth rates could enhance our understanding of the factors influencing demographic trends.
+Further exploration of these topics could contribute to a deeper understanding of the broader societal context surrounding birth rate trends in the United States
 

### 2024-02-12 outputs/paper/paper.qmd
update no times new roman

@@
        caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
-  theme_classic() +
-  theme(text = element_text(family = "Times New Roman"))
+  theme_classic()
 
@@
        title = NULL) +
-  theme_classic() +
-  theme(text = element_text(family = "Times New Roman"))
+  theme_classic()
 

### 2024-02-13 outputs/paper/paper.qmd
Final update

@@
 2016) for creating graphical representations.
-Additionally, the gridExtra package was used to enhance the layout and presentation of the figures (Auguie 2017).
+Additionally, the gridExtra package was used to enhance the layout and presentation of the figures (Murrell 2021).
 
-The paper proceeds by first presenting the trend in US birth rates over the period 1980 to 2020, accompanied by relevant factors (CDC).
-It then delves into variations in birth rates by age group and race/ethnicity of mothers, correlating these trends with education, socio-economic factors, and financial stability.
+The paper proceeds by first presenting the trend in US birth rates over the period 1980 to 2020.
+It then delves into variations in birth rates by age group and race and ethnicity of mothers, correlating these trends with education, socio-economic factors, and financial stability.
 The replication of Kearney, Levine, and Pardue's study focuses on answering the following research questions:
@@
 
-The data on the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021; Hamilton and Osterman 2021).
+The data on the trend in US Birth Rates, measured as births per 1,000 women aged 15-44 from 1980 to 2020, was sourced from the CDC Vital Statistics Births Reports for the years 2015, 2019, and 2020 (Martin and Mathews 2017; Martin and Driscoll 2021).
 Additionally, trends in Birth Rates by Population Subgroup, specifically birth rates categorized by age group, race, and ethnicity, were gathered from the CDC Vital Statistics Births Reports.
@@
 
-The de facto population of women of reproductive age spans from 15 to 49 years in a country (WHO 2024).
+The de facto population of women of reproductive age spans from 15 to 49 years in a country (World Health Organization 2024).
 The average age of the final menstrual period (menopause) is 51, and for women in their 40s, pregnancy is rare but not impossible (ASRM 2012).
@@
      family = "Times") +
-theme(axis.title.y = element_text(size = 12, family = "Times")) +
+theme(axis.title.y = element_text(size = 4, family = "Times")) +
   theme_classic() +

### 2024-02-15 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -8,7 +8,7 @@
 # Pre-requisites: Need to download the replication package
 
 #### Workspace setup ####
-install.packages(c("knitr", "lubridate", "tidyverse"))
+# install.packages(c("knitr", "lubridate", "tidyverse"))
 
 # See a list of available packages
 library(knitr)

### 2024-02-15 scripts/03-test_data.R
Update 03-test_data.R

@@ -11,16 +11,16 @@
 
 
 #### Workspace setup ####
-install.packages("ggplot2")
+# install.packages("ggplot2")
 library(ggplot2)
 
-install.packages("dplyr")
+# install.packages("dplyr")
 library(dplyr)
 
-install.packages("tidyr")
+# install.packages("tidyr")
 library(tidyr)
 
-install.packages("usmap")
+# install.packages("usmap")
 library(usmap)
 
 #### Test data ####

### 2024-02-15 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -2,18 +2,7 @@
 # Purpose: To sort out relevant necessary data from the data set 
 # ~not sure whether I should haave this script tho
 # Author: Chay Park
-# Date: 4 February 2024
+# Date: 15 February 2024
 # Contact: chay.park@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to download the replication package
-
-#### Workspace setup ####
-# install.packages(c("knitr", "lubridate", "tidyverse"))
-
-# See a list of available packages
-library(knitr)
-library(lubridate)
-library(tidyverse)
-
-#### Clean data ####
-

### 2024-02-15 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -1,8 +1,7 @@
 #### Preamble ####
-# Purpose: To sort out relevant necessary data from the data set 
-# ~not sure whether I should haave this script tho
 # Author: Chay Park
 # Date: 15 February 2024
 # Contact: chay.park@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Need to download the replication package
+# Pre-requisites: Need to download the replication package, follow the 01-download_data.R steps, 
+# the paper uses the dataset from the original csv file without modification. 

### 2024-02-15 outputs/paper/paper.qmd
Update paper.qmd

@@
   - Chay Park
-thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Social-Science-Study."
+thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Social-Science-Study.
+A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-srs6-t802""
 date: today

### 2024-02-15 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,8 +1,7 @@
 #### Preamble ####
 # Purpose: The purpose of this script is to simulate and visualize the trend in birth rates in the United States from the year 1980 
 # to 2020. The script generates tables and graphs to illustrate the overall trend in US birth rates, as well as the birth rates among 
-# various subgroups such as different age groups, ethnicities, educational levels, and marital statuses. 
-# Additionally, the script creates a state-level map to visualize the change in birth rates across different states over the specified period.
+# various subgroups such as different age groups and ethnicities
 # Author: Chay Park
 # Date: 3 February 2024 
 # Contact: chay.park@mail.utoronto.ca

### 2024-02-15 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,53 +1,71 @@
 #### Preamble ####
-# Purpose: The purpose of this script is to simulate and visualize the trend in birth rates in the United States from the year 1980 
-# to 2020. The script generates tables and graphs to illustrate the overall trend in US birth rates, as well as the birth rates among 
-# various subgroups such as different age groups and ethnicities
+# Purpose: In the provided code, I simulate the generation of two sample datasets 
+# representing birth rates across different demographics over multiple years. 
+# The goal is to create comprehensive datasets with various demographic categories, 
+# filling in missing values with appropriate defaults, and then perform tests to 
+# ensure non-zero total counts, absence of missing values, and inclusion of all combinations.
 # Author: Chay Park
-# Date: 3 February 2024 
+# Date: 15 February 2024 
 # Contact: chay.park@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: R Programming Environment, R Packages, Simulated Birth Rate Data, Simulated Subgroup Data, State-level birth rate change data
 
 #### Workspace setup ####
-# Load required libraries
-library(ggplot2)
+
+# Load necessary packages
+if (!requireNamespace("dplyr", quietly = TRUE)) {
+  install.packages("dplyr")
+}
 library(dplyr)
-library(tidyr)
 
-#### Simulate data ####
-# Simulated birth rate data from 1980 to 2020
-set.seed(123) # for reproducibility
-birth_data <- data.frame(
-  year = 1980:2020,
-  births = rnorm(41, mean = 1000, sd = 100)
+# Set seed for reproducibility
+set.seed(123)
+
+#### Dataset Simulation ####
+
+# Simulate creating the first dataset
+years <- 2000:2022
+brate_all <- runif(length(years), min = 5, max = 20)
+
+data_first <- data.frame(
+  year = years,
+  brate_all = brate_all
 )
 
-# Generate subgroup data
-subgroup_data <- data.frame(
-  year = rep(1980:2020, each = 6),
-  subgroup = rep(c("brate_1519", "brate_2024", "brate_2529", "brate_hisp", "brate_college", "brate_married"), times = 41),
-  births = rnorm(41*6, mean = 500, sd = 50)
+# Simulate creating the second dataset
+brate_1519 <- runif(length(years), min = 2, max = 10)
+brate_3034 <- runif(length(years), min = 5, max = 15)
[…]

### 2024-02-15 scripts/00-simulate_data.R
Update the simulate_data R script's date and purpose

@@ -1,10 +1,9 @@
 #### Preamble ####
 # Purpose: The purpose of this script is to simulate and visualize the trend in birth rates in the United States from the year 1980 
 # to 2020. The script generates tables and graphs to illustrate the overall trend in US birth rates, as well as the birth rates among 
-# various subgroups such as different age groups, ethnicities, educational levels, and marital statuses. 
-# Additionally, the script creates a state-level map to visualize the change in birth rates across different states over the specified period.
+# various subgroups such as different age groups and race and ethnicities. 
 # Author: Chay Park
-# Date: 3 February 2024 
+# Date: 15 February 2024 
 # Contact: chay.park@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: R Programming Environment, R Packages, Simulated Birth Rate Data, Simulated Subgroup Data, State-level birth rate change data

### 2024-02-15 scripts/03-test_data.R
Rewrite the 03-test_data.R script

@@ -1,35 +1,64 @@
 #### Preamble ####
-# Purpose: The test codes aim to verify the functionality and characteristics of the birth rate dataset created earlier. 
-# They serve to validate the data, assess its distribution, and explore potential relationships between variables.
+# Purpose: In the provided code, I run tests on real datasets representing birth rates across different demographics over multiple years.
 # Author: Chay Park
-# Date: 4 February 2024
+# Date: 15 February 2024
 # Contact: chay.park@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The test codes assume that the birth rate dataset (named birth_data) is available in the R environment and 
 # contains the necessary variables (year and births). Additionally, the R environment with required packages 
-# (e.g., ggplot2, dplyr, tidyr, usmap) should be set up and loaded for executing the test codes successfully.
+# (e.g., ggplot2, dplyr, tidyr) should be set up and loaded for executing the test codes successfully.
 
+
+#### Preamble ####
 
 #### Workspace setup ####
-# install.packages("ggplot2")
-library(ggplot2)
 
-# install.packages("dplyr")
+# Load necessary packages
+if (!requireNamespace("dplyr", quietly = TRUE)) {
+  install.packages("dplyr")
+}
 library(dplyr)
 
-# install.packages("tidyr")
-library(tidyr)
+#### Load real datasets ####
 
-# install.packages("usmap")
-library(usmap)
+# Read the data from CSV file
+fig_1 <- read.csv(file = here::here("inputs/data/fig_1.csv"))
 
-#### Test data ####
-# 1. Test code to check the summary statistics of the birth rate data:
-summary(birth_data)
+# Read the data from CSV file
+figs_2a_2b <- read.csv(file = here::here("inputs/data/fig_1.csv"))
 
-# 2. Test code to visualize the distribution of birth rates over the years:
-hist(birth_data$births, main = "Distribution of Birth Rates", xlab = "Births per 1000 women age 15-44", ylab = "Frequency")
+#### Test codes ####
 
-# 3. Test code to calculate the correlation between birth rates and year:
[…]

### 2024-02-15 outputs/paper/US_birth_rates.qmd
Rename outputs/US_birth_rates.qmd to outputs/paper/US_birth_rates.qmd

--- 2024-02-15 outputs/paper/paper.qmd
@@
 thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Social-Science-Study.
-A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-srs6-t802""
+A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-55pn-6h10"
 date: today
 date-format: long
-abstract: "This paper examines the notable decrease in US birth rates from 2007 to 2020, replicating the study conducted by Kearney, Levine, and Pardue (2022).
-Through an analysis of trends spanning from 1980 to 2020, we investigate variations in birth rates among different age groups and disparities based on the race and ethnicity of mothers.
-While initially attributing the decline to the Great Recession, our analysis uncovers additional factors contributing to this trend.
-Furthermore, our exploration delves into socio-cultural aspects, offering insights for policymakers and researchers interested in population trends and their implications."
+abstract: "This paper replicates Kearney, Levine, and Pardue's (2022) study on the notable decrease in US birth rates from 2007 to 2020.
+Analyzing data spanning from 1980 to 2020, we examine variations in birth rates among different age groups and disparities based on the race and ethnicity of mothers.
+Our analysis uncovers additional factors contributing to this trend beyond the initial attribution to the Great Recession.
+Specifically, socio-economic factors such as contraceptive use, cultural norms, and financial stability play key roles in the observed decline in birth rates."
 format: pdf
@@
 
-According to the Centers for Disease Control and Prevention, the birth rate is defined as "the number of live births per 1,000 population." This rate is calculated by dividing the number of live births in a population in a year by the midyear resident population.
+According to the Centers for Disease Control and Prevention (CDC), the birth rate is defined as "the number of live births per 1,000 population." This rate is calculated by dividing the number of live births in a population in a year by the midyear resident population.
 Specifically, for census years, rates are calculated based on the unrounded census counts of the resident population as of April 1st.
-From an economic perspective, birth rates are an important factor that determines the growth of a country.
+
+From an economic perspective, birth rates are a vital determinant of a country's growth.
 Recent years have seen widespread concern over declining birth rates globally, prompting nations to explore strategies to address this issue.
-For instance, Germany's birth rate in 2024 is 9.3, which is lower than Canada's 10.0 and the US's 12.0 (Macrotrends 2024).
-Germany is taking action by actively recruiting skilled workers from abroad (Grunau 2023) and reforming its citizenship law to allow dual citizenship (Knight 2024) as some measures to combat declining birth rates and overall population.
-Furthermore, birth rates serve as an indicator of replacement-level fertility, essential for maintaining population stability in developed nations (Craig 1994).
+For example, in 2024, Germany's birth rate is reported as 9.3, lower than Canada's 10.0 and the US's 12.0 (Macrotrends 2024).
+Germany is taking action by actively recruiting skilled workers from abroad (Grunau 2023) and reforming its citizenship law to allow dual citizenship (Knight 2024) as measures to combat declining birth rates and overall population decline.
+
+Furthermore, birth rates serve as an indicator of replacement-level fertility, essential for maintaining population stability in developed nations.
+Replacement-level fertility is the level of fertility at which each generation is exactly replaced by the next generation without any population growth or decline.
+It's typically estimated to be around 2.1 children per woman in developed countries, accounting for factors like infant mortality and individuals who don't reproduce.
+When birth rates fall below this threshold, it can lead to population decline, aging populations, and potential economic challenges.
+
+The estimand, or the target of estimation, in this context is the birth rate itself – specifically, the average number of live births per 1,000 population over a specified period.
+Understanding and accurately estimating this estimand is crucial for assessing population dynamics and informing policy decisions aimed at addressing declining birth rates.
 
@@
 
-This paper utilized a replication package of the original study and employed a cleaned data CSV file to reproduce the related figures and tables.
-While the original code in the replication package was written in Stata, this paper utilized R (R Core Team 2020) for analysis, leveraging packages such as ggplot2 (Wickham et al.
+This paper utilizes a replication package of the original study and employs a cleaned data CSV file to reproduce the related figures and tables.
+While the original code in the replication package was written in Stata, this paper utilizes R (R Core Team 2020) for analysis, leveraging packages such as ggplot2 (Wickham et al.
 2016) for creating graphical representations.
@@
 The paper proceeds by first presenting the trend in US birth rates over the period 1980 to 2020.
-It then delves into variations in birth rates by age group and race and ethnicity of mothers, correlating these trends with education, socio-economic factors, and financial stability.
+It then delves into variations in birth rates by age group and race/ethnicity of mothers, correlating these trends with education, socio-economic factors, and financial stability.
 The replication of Kearney, Levine, and Pardue's study focuses on answering the following research questions:
@@
 
-Regarding race and ethnicity, "white" refers to the non-Hispanic Caucasian population group, "black" refers to the non-Hispanic African American group, and "Hispanic" refers to people with Hispanic ethnicity, mainly from South America, for brevity.
+In this paper, data validation is conducted using an R script file named '03-test_data' to ensure the accuracy of the dataset obtained directly from the original package.
+The paper utilizes both 'fig_1' and 'fig_2a_2b' CSV files.
+
+In the dataset, 'brate_all' represents the birth rates of all women aged 15-44 in the US.
+'Brate' is an abbreviation for 'birth rates'.
+For example, 'brate_1519' corresponds to the birth rate data for women aged 15 to 19 years old, irrespective of race and ethnicity.
+Similarly, 'brate_whitenh' represents the birth rates of white, non-Hispanic women, regardless of age.
+
+Regarding race and ethnicity, "white" refers to the non-Hispanic Caucasian population group, "black" refers to the non-Hispanic African American group, and "Hispanic" refers to individuals with Hispanic ethnicity, primarily from South America, for brevity.
 
@@
 # Read the data from CSV file
-data <- read.csv(file = here::here("outputs/data/fig_1.csv"))
+data <- read.csv(file = here::here("inputs/data/fig_1.csv"))
 
@@
 #| label: fig-polscatter-graph
-#| fig-cap: "Trend in US Birth Rates during the presidential mandate"
+#| fig-cap: "Trend in US Birth Rates During the Presidential Mandate"
 
@@
 # Read the data from CSV file
-data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))
+data <- read.csv(file = here::here("inputs/data/figs_2a_2b.csv"))
 
@@
   labs(x = NULL, 
-     y = 'Births per 1,000 women in relevant population subgroup',
+     y = 'Births rates in relevant population subgroup',
      title = "A: Five-year age group",
@@
 Understanding the potential impact of immigration laws and campaign promises during presidential mandates on birth rates could enhance our understanding of the factors influencing demographic trends.
-Further exploration of these topics could contribute to a deeper understanding of the broader societal context surrounding birth rate trends in the United States
+Further exploration of these topics could contribute to a deeper understanding of the broader societal context surrounding birth rate trends in the United States.
 

### 2024-02-15 scripts/00-simulate_data.R
Delete the paper.pdf and paper.qmd file and edit simulate_data.R file

@@ -1,15 +1,7 @@
 #### Preamble ####
-<<<<<<< HEAD
 # Purpose: The purpose of this script is to simulate and visualize the trend in birth rates in the United States from the year 1980 
 # to 2020. The script generates tables and graphs to illustrate the overall trend in US birth rates, as well as the birth rates among 
 # various subgroups such as different age groups and race and ethnicities. 
-=======
-# Purpose: In the provided code, I simulate the generation of two sample datasets 
-# representing birth rates across different demographics over multiple years. 
-# The goal is to create comprehensive datasets with various demographic categories, 
-# filling in missing values with appropriate defaults, and then perform tests to 
-# ensure non-zero total counts, absence of missing values, and inclusion of all combinations.
->>>>>>> a58e1fe82c36a9fb83b6a5c5cc440fde66db2620
 # Author: Chay Park
 # Date: 15 February 2024 
 # Contact: chay.park@mail.utoronto.ca
